---
title: "Issue found"
author: "Miguel Agenjo Montejo"
format: html
editor: visual
---

## Set up

```{r}
#| message: false
#| warning: false

# library(ineapir) (won't use it, too slow)
library(spanishoddata)
library(dplyr)
library(stringr)
library(tidyr)
library(readr)
library(explore)
library(ggplot2)
library(patchwork)
library(mapSpain)
library(sf)
library(RColorBrewer)

# Set the spanishoddata download folder:
spod_set_data_dir("../data/mitma")
```

## Data preprocessing

### Download income and population data from INE

```{r}
# Download income and population data from INE:

# We save the downloaded data to save time in new sessions
# income_adrh_raw <- read.csv2("https://ine.es/jaxiT3/files/t/es/csv_bdsc/30824.csv?nocab=1")
# saveRDS(income_adrh_raw, "income_adrh_raw.rds")
income_adrh_raw <- readRDS("income_adrh_raw.rds")

income_adrh_raw <- as_tibble(income_adrh_raw)

# We save the downloaded data to save time in new sessions
# dem_adrh_raw <- read.csv2("https://ine.es/jaxiT3/files/t/es/csv_bdsc/30832.csv?nocab=1")
# saveRDS(dem_adrh_raw, "dem_adrh_raw.rds")
dem_adrh_raw <- readRDS("dem_adrh_raw.rds")

dem_adrh_raw <- as_tibble(dem_adrh_raw)
```

### Clean INE data

#### Income dataset:

```{r}
glimpse(income_adrh_raw)
describe(income_adrh_raw)
```

```{r}
# We rename the columns, fix column types, convert Spanish decimal separator
# "," into English decimal separator ".", and convert blank spaces into NAs:
income_adrh <- income_adrh_raw |>
  mutate(
    "municipality" = na_if(Municipios, ""),
    "district" = na_if(Distritos, ""),
    "section" = na_if(Secciones, ""),
    "income_indicator" = Indicadores.de.renta.media,
    "year" = Periodo,
    "total" = parse_number(
      Total,
      na = c("", "."),
      locale = locale(grouping_mark = ".", decimal_mark = ",")
    )
  ) |>
  # Select to keep only the clean columns, not the raw ones:
  select(municipality, district, section, income_indicator, year, total)

# Pivot the income indicators wider to get an indicator per column:
income_adrh <- income_adrh |> pivot_wider(
  names_from = income_indicator,
  values_from = total)

# Rename the income indicators:
income_adrh <- income_adrh |> rename(
  "mean_net_income_person" = `Renta neta media por persona`,
  "mean_net_income_household" = `Renta neta media por hogar`,
  "mean_income_per_consumption_unit" = `Media de la renta por unidad de consumo`,
  "median_income_per_consumption_unit" = `Mediana de la renta por unidad de consumo`,
  "mean_gross_income_person" = `Renta bruta media por persona`,
  "mean_gross_income_household" = `Renta bruta media por hogar`
)

# Remove raw data from the environment:
rm(income_adrh_raw)


# Check the result:
glimpse(income_adrh)
describe(income_adrh)
```

We need to keep in mind that we have a relatively small percentage of NAs in the income variables. Let's check what the missings are at the section level, which is what we will actually use:

```{r}
# Keep only the section-level rows (the ones we will actually use):
income_adrh |>
  filter(!is.na(section)) |>
  describe()

# Let's check if the NA's are equally distributed across years:

# And check the missing for each year:
income_adrh |>
  filter(!is.na(section)) |>
  group_by(year) |>
  summarise(
    sections = n(),
    # As we will show later, this is the relevant variable:
    na_net_income_per_person = sum(is.na(mean_net_income_person)), 
    pct_net_income_per_person = 100 * mean(is.na(mean_net_income_person))
  )
```

The missing percentage in is under 2% for the years relevant to us.

#### Demographic dataset:

```{r}
glimpse(dem_adrh_raw)
describe(dem_adrh_raw)
```

```{r}
# We rename the columns, fix column types, convert Spanish decimal separator
# "," into English decimal separator ".", and convert blank spaces into NAs:
dem_adrh <- dem_adrh_raw |>
  mutate(
    "municipality" = na_if(Municipios, ""),
    "district" = na_if(Distritos, ""),
    "section" = na_if(Secciones, ""),
    "demographic_indicator" = Indicadores.demográficos,
    "year" = Periodo,
    "total" = parse_number(
      Total,
      na = c("", "."),
      locale = locale(grouping_mark = ".", decimal_mark = ",")
    )
  ) |>
  # Select to keep only the clean columns, not the raw ones:
  select(municipality, district, section, demographic_indicator, year, total)

# Pivot the demographic indicators wider to get an indicator per column:
dem_adrh <- dem_adrh |> pivot_wider(
  names_from = demographic_indicator,
  values_from = total
)

# Rename the demographic indicators:
dem_adrh <- dem_adrh |> rename(
  "mean_age" = `Edad media de la población`,
  "pct_under_18" = `Porcentaje de población menor de 18 años`,
  "pct_65_and_over" = `Porcentaje de población de 65 y más años`,
  "mean_household_size" = `Tamaño medio del hogar`,
  "pct_single_person_households" = `Porcentaje de hogares unipersonales`,
  "population" = `Población`,
  "pct_spanish_population" = `Porcentaje de población española`
)


# Remove raw data from the environment:
rm(dem_adrh_raw)


# Check the result:
glimpse(dem_adrh)
describe(dem_adrh)
```

Let's check what NA percentage for population is at the section level:

```{r}
# Keep only the section-level rows (the ones we will actually use):
dem_adrh |>
  filter(!is.na(section)) |>
  describe()

# Let's check if the NA's are equally distributed across years:

# And check the missing for each year:
dem_adrh |>
  filter(!is.na(section)) |>
  group_by(year) |>
  summarise(
    sections = n(),
    na_population = sum(is.na(population)),
    pct_na_population = 100 * mean(is.na(population))
  )
```

The NA percentage in the population variable is under 2% in our study period.

Join both tibbles (we need the population of each section from the demographic dataset, which is not available in the income dataset).

```{r}
# Join the tibbles:
adrh <- dem_adrh |>
  # We keep just the join key columns and the population variable we need:
  select(municipality, district, section, year, population) |> 
  # And join:
  left_join(
    income_adrh,
    by = join_by(municipality, district, section, year),
    keep = FALSE
  )

# We save the downloaded data to save time in new sessions
# ine_geojson_2023 <- st_read(url_ine_geojson_2023)
# saveRDS(ine_geojson_2023, "ine_geojson_2023.rds")
ine_geojson_2023 <- readRDS("ine_geojson_2023.rds")

# Change column names:
ine_geojson_2023 <- ine_geojson_2023 |>
  rename(
    "prov_code"  = CPRO,
    "muni_code" = CMUN,   # 3-digit, unique within province
    "prov_muni_code" = CUMUN,  # 5-digit (province + municipality)
    "alt_prov_mun_code" = CLAU2,  # alternate 5-digit municipality
    "dist_code" = CDIS,   # 2-digit, unique within municipality
    "prov_muni_dist_code" = CUDIS,  # 7-digit unified
    "sect_code" = CSEC,   # 3-digit, unique within district
    "prov_muni_dist_sect_code" = CUSEC,  # 10-digit unified
    "autonomous_community_code" = CCA,
    "province_name" = NPRO,
    "autonomous_community_name" = NCA,
    "municipality_name" = NMUN,
    "nuts_level0_code" = CNUT0,
    "nuts_level1_code" = CNUT1,
    "nuts_level2_code" = CNUT2,
    "nuts_level3_code" = CNUT3,
    "type" = TIPO
  )
```

### MITMS data

With all the INE data ready, we download MITMS data for a random day (first of June 2023), a normal working thursday.

```{r}
# Only downloads the first time, after that takes it from the folder we defined
# in the first chunk of the .qmd
od_districts <- spod_get_zones("dist", ver = 2)

od_data_20230601 <- spod_get(
  type  = "od",
  zones = "districts",
  dates = "2023-06-01")
```

## Context on the issue found

While working on other things, I detected an issue with the MITMS od database that is made available through the `spanishoddata` package.

### Household income vs person income, net or gross?

First of all, there is a lack of information in the documentation of the R package. According to the official codebook, available [here](https://ropenspain.github.io/spanishoddata/articles/v2-2022-onwards-mitma-data-codebook), the income variable is:

> The income group of individuals making the trips in n_trips. Categorized into \<10, 10-15, and \>15 (thousands of euros per year). The income for each individual is assigned based on the mean census tract income per person (data source is INE Household income distribution map).

This doesn't clarify wether we are using net or gross income. Moreover, an official document by the Ministerio de Transporte, available [here](https://www.transportes.gob.es/recursos_mfom/paginabasica/recursos/a3_informe_metodologico_estudio_movilidad_mitms_v8.pdf), reads:

> Segmentación por perfil sociodemográfico: se segmentan los viajes de acuerdo a las siguientes características de los viajeros: [...] Nivel de renta media del hogar, según las siguientes franjas en miles de euros: \< 10, entre 10 y 15 y \>15.

That is, it states that the income is categorized according to average household income, which contradicts the official documentation (stating "per person"), and doesn't say anything about whether the variable is net or gross income.

### Settling what income we are talking about

An exploration of the datasets allows us to settle this question, since the only income variable that makes the 10.000 and 15.000 euros thresholds sensible is mean net individual income. As shown below.

```{r}
# Plot distributions of income for the year 2023:
plot_income_distribution <- function(tibble_name, variable_name, year_chosen) {
  tibble_name |>
    # Select only sections and for the chosen year:
    filter(!is.na(section), year == year_chosen) |>
    ggplot(aes(x = .data[[variable_name]])) + # We'll pas it in quotes
    geom_histogram(fill = "steelblue", bins = 50) +
    # Add verticla lines at the 10000 and 15000 thresholds used by MITMS
    geom_vline(xintercept = 10000, color = "gray", linetype = "dashed") +
    geom_vline(xintercept = 15000, color = "gray", linetype = "dashed") +
    labs(x = variable_name, y = "Nº de secciones") +
    theme_minimal()
}

plot_income_distribution(adrh, "mean_net_income_person", 2023) +
plot_income_distribution(adrh, "mean_gross_income_person", 2023) +
plot_income_distribution(adrh, "mean_net_income_household", 2023) +
plot_income_distribution(adrh, "mean_gross_income_household", 2023)
```

### The core of the issue

#### Area overlap

We take a single zone from the MITMS database To prove there are no differences regarding geographical scope (this should be the case according to the documentation, but we confirm it anyway to be completely sure).The chosen zone is 287920, which corresponds to the district of San Blas-Canillejas, in the city of Madrid.

We plot the geometry of this zone from the MITMS database against the corresponding sections provided by INE.

```{r}
sanblas_mitma <- od_districts |>
  filter(id == "2807920")

sanblas_ine <- ine_geojson_2023 |>
  filter(str_starts(prov_muni_dist_sect_code, "2807920"))

# They don't use the same coordinate reference system (CRS), we unify it:
sanblas_mitma <- st_transform(sanblas_mitma, st_crs(sanblas_ine))

# Plot:
ggplot() +
  geom_sf(data = sanblas_ine, fill = "steelblue", color = "white",
          linewidth = 0.2) +
  geom_sf(data = sanblas_mitma, fill = NA, color = "darkred", linewidth = 0.8) +
  labs(title = "San Blas-Canillejas",
       subtitle = "Blue: INE, red outline: MITMS") +
  theme_minimal()
```

We confirm they perfectly overlap, no issue regarding the geometries or the sections that form a zone in the MITMS dataset.

#### The contradiction regarding income

Here lies the core of the issue we found. According to the `spanishoddata` documentation, as we showed above, the income variable "is assigned based on the mean census tract income per person (data source is INE Household income distribution map)". This means that we should see the `<10` label assigned to trips made by people residing in census sections with an average net individual yearly income under 10 thousand euros, and so on for the `10-15` and `>15` labels.

We check what the actual distribution looks like for a random day in 2023:

```{r}
trips_by_bracket <- od_data_20230601 |>
  filter(id_origin == "2807920", activity_origin == "home") |>
  rename("income_bracket" = income) |>
  group_by(income_bracket) |>
  summarise(trips = sum(n_trips, na.rm = TRUE)) |>
  collect()

trips_by_bracket
```

As we can see above, according to the MITMS data there were 177,821 trips made by people in the 10-15 thousand euros income bracket, and only 14 trips made by people with an income over 15 thousand euros. No trips by people with income under 10 thousand euros are recorded. This contradicts the data from INE (which supposedly the MITMS is using to assign the income variable).

```{r}
income_levels <- c("<10", "10-15", ">15")

people_by_bracket <- adrh |>
  filter(str_starts(section, "2807920"), year == 2023) |>
  mutate(
    "income_bracket" = case_when(
      mean_net_income_person > 15000 ~ ">15",
      mean_net_income_person >= 10000 & mean_net_income_person <= 15000 ~ "10-15",
      mean_net_income_person < 10000 ~ "<10")) |> 
  filter(!is.na(income_bracket)) |> 
  mutate("income_bracket" = factor(income_bracket, levels = income_levels)) |> 
  group_by(income_bracket) |>
  summarise(
    sections = n(),
    population = sum(population),
  ) 

people_by_bracket
```

The conclussion is clear, but let's plot it side by side to make sure we deliver our point:

```{r}
plot_trips <- ggplot(trips_by_bracket,
                     aes(x = income_bracket, y = trips)) +
  geom_col(fill = "steelblue") +
  scale_x_discrete(drop = FALSE) +
  labs(x = "", y = "Trips",
       title = "MITMS: trips",
       subtitle = "Home-origin trips, 2023-06-01") +
  theme_minimal() +
  theme(plot.subtitle = element_text(size = 7))

plot_people <- ggplot(people_by_bracket,
                      aes(x = income_bracket, y = population)) +
  geom_col(fill = "steelblue") +
  scale_x_discrete(drop = FALSE) +
  labs(x = "", y = "Población",
       title = "INE: people",
       subtitle = "By mean net per person income of their census section, 2023") +
  theme_minimal() +
  theme(plot.subtitle = element_text(size = 7))

plot_trips + plot_people
```

It is very unlikely that the almost 6000 people living in a census tract with an income of under 10000 don't produce any trips recorded by the telephone operator. There should be some recorded trips in the `<10` bracket. More strikingly, bracket with the most people doesn't seem to produce any trips. Something is wrong, and given the total amount of trips, it almost is surely a problem with income labeling.

We can repeat the same thing for other districts:

## Reproducing the issue with more districts:

We now repeat these same three steps for every district, so we wrap them in a helper function, `ids` is a vector because one INE district can be split into several MITMS zones (Pozuelo de Alarcón, for instance, is both `2811501` and `2811502`); `str_starts()` does not take a vector of prefixes as "starts with any of these", so we collapse them into a single regex alternation.

```{r}
compare_trips_and_people <- function(od_data, adrh_data, ids, year_chosen, label) {

  income_levels <- c("<10", "10-15", ">15")
  section_prefix <- str_c(ids, collapse = "|")
  
  # Take date for the subtitle from the od data:
  date_subtitle <- od_data |>
    distinct(date) |> 
    collect() |> 
    pull(date)

  # MITMS: trips by income bracket
  trips_by_bracket <- od_data |>
    filter(id_origin %in% ids, activity_origin == "home") |>
    group_by(income) |>
    summarise(trips = sum(n_trips, na.rm = TRUE)) |>
    collect() |>
    mutate("income_bracket" = factor(income, levels = income_levels))

  # INE: sections and people by the income bracket of their census section
  people_by_bracket <- adrh_data |>
    filter(str_starts(section, section_prefix), year == year_chosen) |>
    mutate(
      "income_bracket" = case_when(
        mean_net_income_person > 15000 ~ ">15",
        mean_net_income_person >= 10000 & mean_net_income_person <= 15000 ~ "10-15",
        mean_net_income_person < 10000 ~ "<10")) |>
    filter(!is.na(income_bracket)) |> # Sections with no income data
    group_by(income_bracket) |>
    summarise(
      sections = n(),
      population = sum(population),
    ) |>
    mutate("income_bracket" = factor(income_bracket, levels = income_levels))

  # Trips by bracket plot:
  plot_trips <- ggplot(trips_by_bracket,
                       aes(x = income_bracket, y = trips)) +
    geom_col(fill = "steelblue") +
    scale_x_discrete(drop = FALSE) +
    labs(x = "", y = "Trips",
         title = paste(label, "- MITMS: trips"),
         subtitle = paste("Home-origin trips,", date_subtitle)) +
    theme_minimal() +
    theme(plot.subtitle = element_text(size = 7))
  
  # People by bracket plot:
  plot_people <- ggplot(people_by_bracket,
                        aes(x = income_bracket, y = population)) +
    geom_col(fill = "steelblue") +
    scale_x_discrete(drop = FALSE) +
    labs(x = "", y = "Población",
         title = paste(label, "- INE: people"),
         subtitle = paste(
           "By mean net per person income of their census section,",
           year_chosen)) +
    theme_minimal() +
    theme(plot.subtitle = element_text(size = 7))

  # SHow tibbles and plot:
  print(trips_by_bracket)
  print(people_by_bracket)
  plot_trips + plot_people
}
```

### District of Salamanca (Madrid)

```{r}
compare_trips_and_people(od_data_20230601, adrh, "2807904", 2023, "Salamanca")
```

Salamanca does not reproduce the issue, and that is informative in itself. MITMS labels every single trip as `>15`, and the INE data agrees: 124 of the 126 census sections of the district are above 15 thousand euros (the other 2 have no income data). The district is socially homogeneous, so a single label for the whole zone is indistinguishable from a correct section by section assignment. In other words, it seems like the issue only becomes visible in districts that are internally mixed.

### Hortaleza (Madrid)

```{r}
compare_trips_and_people(od_data_20230601, adrh, "2807916", 2023, "Hortaleza")
```

Hortaleza does reproduce the issue, almost 30,000 people from 22 census tracts have no trips recorded in the MITMS datset. All trips in the district of hortaleza are assigned the `>15` label.

### Puente de Vallecas (Madrid)

```{r}
compare_trips_and_people(od_data_20230601, adrh, "2807913", 2023, "Puente de Vallecas")
```

The same issue arises once again: not a single trip receives the`<10` label, while INE shows 26 sections with almost 34 thousand people in that income range. We also would intuitively expect a higher number of trips in the `>15` income bracket.

### Sants-Montjuic (Barcelona)

```{r}
compare_trips_and_people(od_data_20230601, adrh, "0801903", 2023, "Sants-Montjuïc")
```

The issue is not specific to Madrid, as we can see above the case of the district of Sants-Montjuic, in Barcelona has a similar issue.

### Seville, district 09

```{r}
compare_trips_and_people(od_data_20230601, adrh, "4109109", 2023, "Seville, district 09")
```

Once again, this time for Seville.

### València, district 12

```{r}
compare_trips_and_people(od_data_20230601, adrh, "4625012", 2023, "València, district 12")
```

Same thing in Valencia.

### Bilbao, district 04

```{r}
compare_trips_and_people(od_data_20230601, adrh, "4802004", 2023, "Bilbao, district 04")
```

And once again, we find the same phenomena in Bilbao.

## Reproducing for year 2022:

We repeat the same comparison for a different, arbitrarily chosen date, this time in 2022, to check that the issue is not specific to the single day (2023-06-01) used above. Let's check the 16th of November 2022, a random working wednesday. Since it is also from a different year, we use the corresponding year data from INE.

```{r}
od_data_20221116 <- spod_get(
  type  = "od",
  zones = "districts",
  dates = "2022-11-16")
```

```{r}
compare_trips_and_people(od_data_20221116, adrh, "2807904", 2022, "Salamanca")
compare_trips_and_people(od_data_20221116, adrh, "2807916", 2022, "Hortaleza")
compare_trips_and_people(od_data_20221116, adrh, "2807913", 2022, "Puente de Vallecas")
compare_trips_and_people(od_data_20221116, adrh, "0801903", 2022, "Sants-Montjuïc")
compare_trips_and_people(od_data_20221116, adrh, "4109109", 2022, "Seville, district 09")
compare_trips_and_people(od_data_20221116, adrh, "4625012", 2022, "València, district 12")
compare_trips_and_people(od_data_20221116, adrh, "4802004", 2022, "Bilbao, district 04")
```

Having established there is an issue regarding the income variable in the MITMS od dataset, let's explore the dimension of the issue.

## Dimension of the issue

### Are the assigned trips correct and some trips missing? Or are there missasignments?

This first question tells us valuable information about the nature of the issue: whether we just have missing data or there is something else (or different) . Actually, we have no way to check if there is missing data other than finding seemingly impossible results, but we have a way to confirm whether missaignment of income brackets is taking place: checking whether MITMS zones with a single census tract assigned income correctly.

We start by getting the income bracket of every census tract, according to INE:

```{r}
tract_brackets <- adrh |>
  filter(!is.na(section), year == 2023) |>
  # Isolate the 10 digit tract code, which is what MITMS uses:
  mutate("section_code" = str_extract(section, "^\\d{10}")) |>
  mutate(
    "ine_bracket" = case_when(
      mean_net_income_person > 15000 ~ ">15",
      mean_net_income_person >= 10000 ~ "10-15",
      mean_net_income_person < 10000 ~ "<10")) |>
  select(section_code, ine_bracket, population)

glimpse(tract_brackets)
```

Then get the MITMS zones made of a single census tract:

```{r}
single_tract_zones <- od_districts |>
  st_drop_geometry() |>
  # MITMS separates tract codes with ";", if there is no ";" it means there is 
  # a single tract:
  filter(str_count(census_sections, ";") == 0) |>
  select(id, census_sections)
```

We now get the trips of those zones, with the income bracket MITMS assigns them:

```{r}
single_tract_trips <- od_data_20230601 |>
  # Keep the trips with one of their two ends at the traveller's home:
  filter(activity_origin == "home" | activity_destination == "home") |>
  # Create a residence zone id:
  mutate(
    "residence_zone_id" = if_else(
      activity_origin == "home", # If a trip starts at home,
      id_origin, # then the zone where that trip starts is the id,
      id_destination)) |>  # otherwise, the destination is home, thus the id.
  # And get the amount of trips by income bracket (should just be one bracket):
  group_by(residence_zone_id, income) |>
  summarise(trips = sum(n_trips, na.rm = TRUE), .groups = "drop") |>
  collect() |>
  mutate("mitma_bracket" = as.character(income)) |>
  # Keep only the single tract zones:
  semi_join(single_tract_zones, by = c("residence_zone_id" = "id"))

# Each zone should get a single bracket:
nrow(single_tract_trips)
n_distinct(single_tract_trips$residence_zone_id)
```

We get 69 rows for 66 zones, so some zones are reporting trips in more than one income bracket (although this shouldn't happen, since there is just one census tract in these zones). We check what are they reporting:

```{r}
single_tract_trips |>
  group_by(residence_zone_id) |>
  # Keep the zones with more than one row, that is, more than one bracket:
  filter(n() > 1) |>
  ungroup() |>
  arrange(residence_zone_id, desc(trips)) |>
  select(residence_zone_id, mitma_bracket, trips)
```

In these three zones the second bracket is residual: 317 trips against 5857, 18 against 6110, and 5 against 5349. This is a small anomaly compared to the issue we are exploring. It is hard to determine why MITMS gives these zones 2 income brackets when INE only reports one.

In any case, we can continue determining how many of these single census section MITMS zones match the income bracket that INE reports.

The next step is to join what INE says with what MITMS says, for each of those zones:

```{r}
single_tract_comparison <- single_tract_zones |>
  left_join(tract_brackets, by = c("census_sections" = "section_code")) |>
  inner_join(single_tract_trips, by = c("id" = "residence_zone_id")) |>
  # Tracts with no income data tell us nothing, we drop them:
  filter(!is.na(ine_bracket)) |>
  select(id, ine_bracket, mitma_bracket, population, trips)

glimpse(single_tract_comparison)
```

And see where they disagree:

```{r}
single_tract_comparison |> 
  filter(ine_bracket != mitma_bracket) |> 
  nrow()
```

There are 30 MITMS zones assigning a bracket the INE data does not support. In 27 of them the only bracket they report is the wrong one; the remaining 3 are the zones we just saw reporting two brackets, where one of the two matches INE and the other does not.

### Multiple-census-tract zones

There is no way to determine exactly how many multiple-census-tract zones are affected, since there's no strict criteria for deciding when a given section is affected. It it is possible for a census tract to produce a low number of trips relative to their population. However, when that number is really low, it becomes implausible. Specially when other income brackets in the same zone are producing a much higher number of trips (after all, there are just 3 income brackets, with the cutoffs near the most dense part of the distribution).

In what follows, we carry out a very conservative test in order approximate the scale of the issue.

We start by finding the MITMS zones (no matter the number of census tracts) in which some income bracket records fewer trips than 10% of the population assigned to that bracket. Again, although strictly speaking this is not impossible, it would require fewer than one in ten people in that bracket making a single trip of at least 500m during the whole day (and, for the trip count here, a trip from home to a different place and the corresponding trip back home would count as two different trips). This already gives us a good proxy to broadly understand the scale of the issue with the data.

We then go a step further and filter for those MITMS zones where some bracket is producing less trips than 10% its population and, at the same time, another income bracket is producing a number of trips at least equal to its population (100%). Of course, we can only do that with the MITMS zones that include multiple census tracts.

#### Brackets with fewer trips than 10% of their population

```{r}
# We create a tibble with one row per zone and census tract, for every zone
# (MITMS separates the tract codes with ";"):
zone_tracts <- od_districts |>
  st_drop_geometry() |>
  filter(!is.na(census_sections)) |>
  select(id, census_sections) |>
  separate_longer_delim(census_sections, ";") |>
  # Remove the spaces:
  mutate("census_sections" = str_trim(census_sections))

# We create a tibble with the population each zone has in each income bracket,
# adding up the tracts that fall in the same bracket:
population_by_zone_bracket <- zone_tracts |>
  left_join(tract_brackets, by = c("census_sections" = "section_code")) |>
  # Tracts with no income data cannot be assigned to a bracket, we remove them:
  filter(!is.na(ine_bracket)) |>
  # Group by zone and income bracket and sum the population:
  group_by(id, ine_bracket) |>
  summarise(population = sum(population, na.rm = TRUE), .groups = "drop")


# We then create a tibble with the trips each MITMS zone produces in each
# income bracket:
trips_by_zone_bracket <- od_data_20230601 |>
  # Keep the trips with one of their two ends at the traveller's home:
  filter(activity_origin == "home" | activity_destination == "home") |>
  # Create a residence zone id:
  mutate(
    "residence_zone_id" = if_else(
      activity_origin == "home", # If a trip starts at home,
      id_origin, # then the zone where that trip starts is the id,
      id_destination)) |>  # otherwise, the destination is home, thus the id.
  group_by(residence_zone_id, income) |>
  summarise(trips = sum(n_trips, na.rm = TRUE), .groups = "drop") |>
  collect() |>
  mutate("mitma_bracket" = as.character(income)) |>
  select(residence_zone_id, mitma_bracket, trips)

# We join the population and trips tibble and compute the ratio of trips by 
# population of each zone and income bracket:
zone_bracket_ratios <- population_by_zone_bracket |>
  # Left join so we keep the brackets MITMS never reports:
  left_join(
    trips_by_zone_bracket,
    by = c("id" = "residence_zone_id", "ine_bracket" = "mitma_bracket")) |>
  mutate(
    # A bracket absent from the MITMS data means zero trips, not missing data:
    "trips" = replace_na(trips, 0),
    "trips_per_person" = trips / population)

# Zones where some bracket produces fewer trips than 10% of its population:
zones_below_threshold <- zone_bracket_ratios |>
  filter(trips_per_person < 0.1) |>
  distinct(id) 

# Number of such zones:
nrow(zones_below_threshold)

# We perform the same filter, but adding a minimum population per zone and
# bracket filter (at least 1000 people in that zone and bracket):
zones_below_threshold_min_pop <- zone_bracket_ratios |>
  filter(trips_per_person < 0.1) |>
  filter(population > 1000) |> 
  distinct(id)

nrow(zones_below_threshold_min_pop)
```

There are 2532 MITMS zones that have at least one income bracket with a number of trips (with origin or destination "home") lower than 10% the population in that zone and income bracket. If make an even more conservative estimation and filter only for those district brackets with more than 1000 population (a highly conservative filter to make sure we are not capturing noise from income brackets with a handful inhabitants), there are still 2371 MITMS zones. That suggests that over half of the zones in the data set are affected.

In the next subsection we create an even more conservative filter keeping, among these zones, those that have another income bracket producing a number of trips at least as big as their population.

#### Zones combining a bracket under 10% with another over 100%

```{r}
# Create a tibble with the zone that have an income bracket producing at least
# as many trips as the population of that bracket:
zones_trips_above_population <- zone_bracket_ratios |>
  filter(trips_per_person >= 1) |>
  distinct(id)

nrow(zones_trips_above_population)

# Join with the 10% threshold tibble, keeping only the zones present in both,
# so that each zone has a bracket under 10% and another one at 100% or above:
zones_both_extremes <- zones_below_threshold |>
  semi_join(zones_trips_above_population, by = "id")

# Number of zones in this new filter:
nrow(zones_both_extremes)

# And the same, requiring the low bracket to hold more than 1000 people, so we
# are not capturing noise from brackets with a handful of inhabitants:
zone_bracket_ratios |>
  filter(trips_per_person < 0.1, population > 1000) |>
  distinct(id) |>
  semi_join(zones_trips_above_population, by = "id") |>
  nrow()
```

There are 2757 MITMS zones with at least one income bracket producing as many trips as the people it holds, or more. Crossing them with the previous filter, 1767 MITMS zones have both things happening at the same time: one income bracket barely producing any trips, and another one producing at least one trip per inhabitant. That is extremely implausible, specially if take into account that we only have three very close income brackets. If we filter only for those where the low bracket includes more than 1000 people to make sure we are not capturing noise, there are still 1632 zones, close to half the comparable zones.

This second filter is stronger than the first one because the comparison is internal to the zone. A bracket producing very few trips could always be explained by the zone as a whole travelling little that day (for example because of bad weather, or holiday taking place). But that explanation cannot hold when a different bracket of the same zone, on the same day, is producing more than one trip per inhabitant.

## The origin of the issue

### The origin of the issue is not the income variable

Below, we run the same test we just ran for `mean_net_income_person` with other income variables to prove that the issue does not arise beacause a different income variable is being used. As can be seen, that is not the case.

The stricter (and shorter) test for just the zones with a single census tract is enough to determine this:

```{r}
# The six income variables the ADRH tables give us:
income_variables <- c(
  "mean_net_income_person", "mean_net_income_household",
  "mean_income_per_consumption_unit", "median_income_per_consumption_unit",
  "mean_gross_income_person", "mean_gross_income_household")

# We create a helper function with the same test we did above to avoid
# repeating the code for each income variable:
test_income_variable <- function(income_variable) {
  variable_brackets <- adrh |>
    filter(!is.na(section), year == 2023) |>
    mutate(
      "section_code" = str_extract(section, "^\\d{10}"),
      "income_value" = .data[[income_variable]],
      "ine_bracket" = case_when(
        income_value > 15000 ~ ">15",
        income_value >= 10000 ~ "10-15",
        income_value < 10000 ~ "<10")) |>
    select(section_code, ine_bracket)

  comparison <- single_tract_zones |>
    left_join(variable_brackets, by = c("census_sections" = "section_code")) |>
    inner_join(single_tract_trips, by = c("id" = "residence_zone_id")) |>
    filter(!is.na(ine_bracket))

  tibble(
    income_variable = income_variable,
    zones_compared = nrow(comparison),
    mismatches = sum(comparison$ine_bracket != comparison$mitma_bracket))
}

# Run it once per income variable:
mismatches_by_variable <- bind_rows(lapply(income_variables, test_income_variable))

mismatches_by_variable
```

The other income variables don't solve the issue, in fact they generate substantially more mismatches.

### The origin of the issue is the year of the INE data

The origin lies in the year of the INE data being used. Instead of using the data for the year the trips belong to, MITMS uses an earlier one. We run the same single-census-tract test against every year the ADRH tables offer:

```{r}
# We create a helper function with the same test we did above, this time
# changing the INE year instead of the income variable:
test_ine_year <- function(ine_year) {
  year_brackets <- adrh |>
    filter(!is.na(section), year == ine_year) |>
    mutate(
      "section_code" = str_extract(section, "^\\d{10}"),
      "ine_bracket" = case_when(
        mean_net_income_person > 15000 ~ ">15",
        mean_net_income_person >= 10000 ~ "10-15",
        mean_net_income_person < 10000 ~ "<10")) |>
    select(section_code, ine_bracket)

  comparison <- single_tract_zones |>
    left_join(year_brackets, by = c("census_sections" = "section_code")) |>
    inner_join(single_tract_trips, by = c("id" = "residence_zone_id")) |>
    filter(!is.na(ine_bracket))

  tibble(
    ine_year = ine_year,
    zones_compared = nrow(comparison),
    mismatches = sum(comparison$ine_bracket != comparison$mitma_bracket))
}

# Run it once per year available in the ADRH tables:
ine_years <- sort(unique(adrh$year))

mismatches_by_year <- bind_rows(lapply(ine_years, test_ine_year))

mismatches_by_year
```

The mismatches fall from 30 with the 2023 data to 3 with the 2020 data, and rise again on both sides of that year. This tells us that the MITMS dataset is using income data from INE's 2020 ADRH, that's why the years closer to 2020 perform better (census section income means change slowly). Note that, as we saw above, the 3 mismatches are due to the MITMS dataset reporting 2 income brackets in some of the single census section zones. Thus, the 2020 net income per person perfectly fits the data reported by MITMS (at least as perfectly as possible).

Knowing which data MITMS is actually using, we can run the threshold test of the previous section again, this time against the 2020 brackets. If the test was sensible and the year explains the issue, far fewer zones should survive the filter:

```{r}
# The income bracket of each census tract, this time according to the 2020 data:
tract_brackets_2020 <- adrh |>
  filter(!is.na(section), year == 2020) |>
  mutate(
    "section_code" = str_extract(section, "^\\d{10}"),
    "ine_bracket" = case_when(
      mean_net_income_person > 15000 ~ ">15",
      mean_net_income_person >= 10000 ~ "10-15",
      mean_net_income_person < 10000 ~ "<10")) |>
  select(section_code, ine_bracket, population)

# The population each zone has in each bracket, exactly as before:
population_by_zone_bracket_2020 <- zone_tracts |>
  left_join(tract_brackets_2020, by = c("census_sections" = "section_code")) |>
  filter(!is.na(ine_bracket)) |>
  group_by(id, ine_bracket) |>
  summarise(population = sum(population, na.rm = TRUE), .groups = "drop")

# The trips are the same ones, only the brackets we compare them against change:
zone_bracket_ratios_2020 <- population_by_zone_bracket_2020 |>
  left_join(
    trips_by_zone_bracket,
    by = c("id" = "residence_zone_id", "ine_bracket" = "mitma_bracket")) |>
  mutate(
    "trips" = replace_na(trips, 0),
    "trips_per_person" = trips / population)

# Zones where some bracket produces fewer trips than 10% of its population,
# and the same filter requiring the bracket to hold more than 1000 people:
zones_below_threshold_2020 <- zone_bracket_ratios_2020 |>
  filter(trips_per_person < 0.1) |>
  distinct(id)

zones_below_threshold_min_pop_2020 <- zone_bracket_ratios_2020 |>
  filter(trips_per_person < 0.1) |>
  filter(population > 1000) |>
  distinct(id)

# Zones compared, and zones in each of the filters:
n_distinct(zone_bracket_ratios_2020$id)
nrow(zones_below_threshold_2020)
nrow(zones_below_threshold_min_pop_2020)
```

The number of affected zones drops a lot, but it is not close to disappearing. Using the 2020 brackets, 1594 of the 3709 comparable zones still hold an income bracket producing fewer trips than 10% of its population, and 1483 of them hold more than 1000 people in that bracket. Against the 2532 and 2371 we got with the 2023 data, the wrong year accounts for roughly a third of the problem, and two thirds of it remain.

A possibility is that the income of each individual is not assigned at the census tract level, but at the MITMS zone level (that is: each individual receives the mean income of the full MITMS zone, instead of the mean of their specific census tract). This doesn't seem to be the spirit of what the documentation of `spanishoddata` states, as we mentioned at the beginning, it reads: “The income for each individual is assigned based on the mean census tract income per person (data source is INE Household income distribution map).” On top of that, this (income being assigned at the MITMS zone level) shouldn't work because many MITMS zones have trips in more than one income trips. However, as we've seen in the zones we manually checked, the "secondary" brackets seem to be residual. Thus, there's some room to believe that the income could have been assigned at the zone level. We check that posibility in the next section.


## Is income assigned at the MITMS zone level?

We test the possibility that income is not assigned based on the individual's census tract but on the MITMS zone where they reside.

```{r}
# First, we create a tibble with the income and population (both from INE) of
# each census section for year 2020:
section_income_2020 <- adrh |>
  filter(!is.na(section), year == 2020) |>
  # The section field carries the code followed by its name:
  mutate("section_code" = str_extract(section, "^\\d{10}")) |>
  select(section_code, population, mean_net_income_person)

# We then attach it to the census sections of each MITMS zone:
zone_section_income_2020 <- zone_tracts |>
  left_join(section_income_2020, by = c("census_sections" = "section_code")) |>
  # A section with no income or no population cannot enter a weighted mean:
  filter(!is.na(mean_net_income_person), !is.na(population), population > 0)

# And calculate the weighted mean income of each MITMS zone and convert it
# into factors with the three MITMS levels (with cuts at 10 and 15):
zone_income <- zone_section_income_2020 |>
  group_by(id) |>
  summarise(
    "zone_income" = sum(mean_net_income_person * population) / sum(population),
    .groups = "drop") |>
  mutate(
    "ine_zone_bracket" = case_when(
      zone_income > 15000 ~ ">15",
      zone_income >= 10000 ~ "10-15",
      zone_income < 10000 ~ "<10"))

# Then create a tibble with the "main" income level of each MITMS zone (from
# trips with origin or destination "home"), keeping the percentage of the 
# zone's trips that falls in it:
zone_main_bracket <- trips_by_zone_bracket |>
  group_by(residence_zone_id) |>
  mutate("main_bracket_share" = trips / sum(trips)) |>
  # The bracket holding most of the zone's trips:
  slice_max(trips, n = 1, with_ties = FALSE) |>
  ungroup() |>
  select("id" = residence_zone_id, mitma_bracket, main_bracket_share)
```


We can now check whether most of the MITMS zones only have a residual amount of trips in their "secondary" brackets. For each zone we take the share of its trips falling in its main income bracket (the one with most trips) and look at how those percentages are distributed across the zones:

```{r}
ggplot(zone_main_bracket, aes(x = main_bracket_share)) +
  geom_histogram(binwidth = 0.02, fill = "steelblue") +
  scale_x_continuous(labels = scales::percent) +
  labs(
    x = "Percentage of the zone's trips in its main income bracket",
    y = "MITMS zones") +
  theme_minimal()
```

```{r}
quantile(zone_main_bracket$main_bracket_share, c(0, 0.1, 0.25, 0.5, 0.75, 1)) 
```

The distribution confirms that most of the zones concentrate the majority of the trips in one income bracket (three fourths of the zones concentrate more than 99.8% of trips in just one bracket; and only the bottom 10% falls below 81.4% of trips concentrated in one income bracket). This suggest that the assignment of trips in secondary brackets is too residual to be the result of having assigned trips at the census level.


```{r}
# We now join both tibbles:
zone_bracket_comparison <- zone_income |>
  inner_join(zone_main_bracket, by = "id") |>
  mutate("matches" = ine_zone_bracket == mitma_bracket)

# Check in how many MITMS zones the "main" income level derived from trips
# matches the mean income level derived from INE:
# Comparable zones (have both income and trips with "home" label):
nrow(zone_bracket_comparison)
# Matches:
sum(zone_bracket_comparison$matches)
# Percentage of matches:
mean(zone_bracket_comparison$matches)
```

It seems like assigning income based on the MITMS zone income (population weighted average from INE census tracts) correctly reproduces MITMS's main bracket in 3399 of the 3700 comparable zones (91.9% of them). This, combined with the fact that the "secondary" brackets seem to have a residual amount of trips, suggests that income data was assigned based on the MITMS zone mean income. 


However, it remains to be explained why 8.1% of the zones don't match what we'd expect if income was assigned at the MITMS zone level.



## Conclusiones

We started by pointing out the contradictions between the official documentation of `spanishoddata` and a document from the MITMS regarding what `income` is in this dataset.


The documentation reads:

> The income group of individuals making the trips in n_trips. Categorized into \<10, 10-15, and \>15 (thousands of euros per year). The income for each individual is assigned based on the mean census tract income per person (data source is INE Household income distribution map).

While the MITMS document reads:

> Segmentación por perfil sociodemográfico: se segmentan los viajes de acuerdo a las siguientes características de los viajeros: [...] Nivel de renta media del hogar, según las siguientes franjas en miles de euros: \< 10, entre 10 y 15 y \>15.

That is, the documentation states that income is *per person* while the MITMS document says it is assigned at the *household* level. Neither of these documents tells us whether it is net or gross income. We established that the variable being used is net income per person (it has the most sensible distribution given what the brackets are, and more decisively, it is the only one that correctly predicts the brackets assigned by MITMS for each specific zone).


In addition, while it is said in the documentation of `spanishoddata` that the source for the income variable is the ADRH map, it is not completely clear whether the data for each individual is assigned at the census tract level or at the MITMS zone level. While the documentation text seems to indicate that the income of each individual is being assigned at the census tract level and only then the data is being aggregated into MITMS zones, our analysis suggests this is not the case (and income is being assigned at the MITMS zone level).

If each individual was assigned the income bracket of their census tract, a MITMS zone made of tracts in different brackets should report trips in all of those brackets, in a proportion we would more or less expect to follow the distribution of the population across them. But that is not what we found. We found that brackets with thousands of inhabitants produce close to no trips (sometimes no trips at all), while almost all trips in each zone fall in one "main" bracket. We saw how over half the zones put every one of their trips in a single income bracket, and three fourths of the zones put more than 99.8% of their trips in one single income bracket. We have to go down to the bottom 10% of the zones to find any that give their main income bracket less than 81.4% of the trips in the zone. In other words, when more than one bracket appears, quite often the secondary brackets are residual, which is the opposite of what we would expect if income was assigned at the census tract level.

We are not completely sure what the assignment mechanism for income was. On the one hand, as we just described, income cannot have been assigned at the census tract level. However, and while assigning the population weighted mean income (of all the census tracts) of each zone does a pretty good job at predicting the income assigned by MITMS, it cannot be just that. If it was just the zone income, we would expect to have just one income bracket for the trips with "home" as origin or destination in each zone (corresponding to the income bracket where the mean income for that MITMS zone falls). And, as we said, a third of the zones do show secondary income brackets; even if residual in the number of trips, they do exist.


Finally, and this is perhaps the most striking part, no document states what's the year of the income data being used. Given this lack of information, the obvious assumption would be that the income variable being used is the one for the year in which the trip happens. However, that is not the case. We took the zones made of a single census tract (where there is no ambiguity about which tract the income should come from) and compared the bracket MITMS reports for their 2023 trips against the bracket each of the available years in the ADRH tables gives them, counting how many match, and the best result is obtained for year 2020. In other words: the income data being used for the trips in 2023 is the one published for the 2020 ADRH. That is a three year gap, enough to change the income bracket a lot of zones fall into, as can be seen below.

```{r}
# The income and population of each census section, now for the two years:
section_income_2020_2023 <- adrh |>
  filter(!is.na(section)) |>
  filter(year %in% c(2020, 2023)) |>
  # The section field carries the code followed by its name, we extract
  # the codes and keep just the relevant variables:
  mutate("section_code" = str_extract(section, "^\\d{10}")) |>
  select(section_code, year, population, mean_net_income_person)

# We now join that to the tibble with the census sections of each MITMS zone:
zone_section_income_2020_2023 <- zone_tracts |>
  # Each section now appears twice, once per year, so every zone-section row
  # of the join comes out duplicated into one row per year:
  left_join(
    section_income_2020_2023,
    by = c("census_sections" = "section_code")) |>
  # A section with no income or no population cannot enter a weighted mean,
  # so we delete them:
  filter(!is.na(mean_net_income_person), !is.na(population), population > 0)

# And calculate the same population weighted zone income as above, but now
# getting one income per zone and year:
zone_income_2020_2023 <- zone_section_income_2020_2023 |>
  group_by(id, year) |>
  summarise(
    "zone_income" = sum(mean_net_income_person * population) / sum(population),
    .groups = "drop") |>
  # As a factor so we can use the year to colour the densities:
  mutate("year" = factor(year))

# Then plot:
ggplot(zone_income_2020_2023, aes(x = zone_income, fill = year)) +
  geom_density(alpha = 0.5, colour = NA) +
  # The two cuts MITMS uses:
  geom_vline(xintercept = c(10000, 15000), linetype = "dotted") +
  # Income in thousands of euros, to read it against the cuts:
  scale_x_continuous(labels = scales::label_number(scale = 1e-3,
                                                   suffix = "k")) +
  scale_fill_manual(values = c("2020" = "steelblue", "2023" = "darkorange")) +
  labs(
    x = "Mean net income per person of the MITMS zone (from INE data)",
    y = "Density",
    fill = "Year") +
  theme_minimal()
```


```{r}
# And count how many zones fall in each bracket each year, cutting the zone
# income at the same 10.000 and 15.000 euros MITMS uses:
zones_per_bracket <- zone_income_2020_2023 |>
  mutate(
    "ine_zone_bracket" = case_when(
      zone_income > 15000 ~ ">15",
      zone_income >= 10000 ~ "10-15",
      zone_income < 10000 ~ "<10")) |>
  count(year, ine_zone_bracket) |>
  # One row per year and one column per bracket, to read it side by side:
  pivot_wider(
    names_from = ine_zone_bracket,
    values_from = n,
    values_fill = 0) |>
  relocate("year", "<10", "10-15", ">15")

zones_per_bracket
```


```{r}
# We now check how many zones are not in the same bracket in both years:
zone_bracket_change <- zone_income_2020_2023 |>
  # Create the brackets:
  mutate(
    "ine_zone_bracket" = case_when(
      zone_income > 15000 ~ ">15",
      zone_income >= 10000 ~ "10-15",
      zone_income < 10000 ~ "<10")) |>
  # Select the relevant variables:
  select(id, year, ine_zone_bracket) |>
  # One column per year, so each zone becomes a single row:
  pivot_wider(
    names_from = year,
    values_from = ine_zone_bracket,
    names_prefix = "bracket_") |>
  # Only the zones with income data in both years can be compared:
  filter(!is.na(bracket_2020), !is.na(bracket_2023)) |>
  # Find the zones for which the bracket changed:
  mutate("changed" = bracket_2020 != bracket_2023) |> 
  # Filter for those that did change:
  filter(changed == TRUE)

# And count them:
nrow(zone_bracket_change)
```


The change is drastic, from 980 zones in the `<10` bracket if we use 2020 data to just over a tenth of that (113) if we use 2023 data. From 371 zones in the `>15` bracket in 2020 to 1188 in 2023. A total of 1663 zones changed which bracket they fall into, which is close to half the zones in the dataset.
