---
title: "Move the canary islands in the MITMS geometries"
author: "Miguel Agenjo Montejo"
format: html
editor: visual
---

## Overview

The geometries provided by the `spanishoddata` R package have the Canary islands in their actual geographical position. In this document we move them using `esp_move_can()` from the `mapSpain` package, and save the resulting data in an .rds file which we will load whenever we need to plot a map.

## Set up

```{r}
#| message: false
#| warning: false

library(dplyr)
library(stringr)
library(sf)
library(spanishoddata)
library(mapSpain)
library(ggplot2)
```

## Load the geometries

```{r}
# Set the data directory:
spod_set_data_dir("../data/mitma/")

# Download the MITMS districts geometries:
mitma_districts <- spod_get_zones(zones = "districts", ver = 2)

# Drop the Portuguese, French, and "external" zones: their ids are "external"
# or start with "PT" or "FR" instead of a digit:
mitma_districts <- mitma_districts |>
  filter(str_starts(id, "\\d"))
```

## Move the Canary Islands with mapSpain

```{r}
# We move the canary islands using esp_move_can() from mapSpain:
canarias <- mitma_districts |>
  filter(str_starts(id, "35|38")) |>  # We keep the Canary Islands only
  esp_move_can(moveCAN = TRUE) # And move them

# Remove the old canary islands and bind the rows of their adjusted geometry:
mitma_districts <- mitma_districts |>
  filter_out(str_starts(id, "35|38")) |>  # Remove the old canary islands
  bind_rows(canarias) # Add the moved canary islands
```

```{r}
# Check it worked:
ggplot(mitma_districts) +
  geom_sf()
```

## Save the geometries

```{r}
# Create the directory to save the file there:
dir.create("../data", showWarnings = FALSE, recursive = TRUE)

# Every map in the thesis reads this file, so that all of them use exactly the
# same geometry, moved the same way:
saveRDS(mitma_districts, "../data/mitma_geometries_moved_can.rds")
```
