---
title: "Estimating the models (terminating trips)"
author: "Miguel Agenjo Montejo"
format: html
editor: visual
---

## Overview and set up

In this document we run the reproduction models, the ones that allow us to reproduce Renninger and Cabrera (2026) and answer RQ1, on the first of the two samples built in `06_build_panels_for_models.qmd`: the trips terminating in the district. The same models are run on the other sample in `07b_estimate_the_models_within_district.qmd`, which also holds the intersectionality models that answer RQ2. This sample is used to run models ABCD for all trips terminating in district i at time t to confirm an unclear point about the sample used by Renninger and Cabrera (2026), as explained in `06_build_panels_for_models.qmd`.

The first four models share the same specification, and differ only in the income variable used (and the subset of trips that variable can be assigned to): model A (full sample of trips, MITMS income) uses all the trips in the sample and the income variable provided by the MITMS dataset; model B (subsample, MITMS income), uses just the subsample of trips with origin home, keeping the income variable provided by the MITMS dataset; model C (subsample, INE income) uses the same subsample of trips but instead of the MITMS income variable, it uses the the income data obtained from INE's ADRH, converted into the same three brackets (with cuts at 10,000 and 15,000 euros) that the MITMS dataset uses; finally, model D (subsample, INE terciles) uses the same subset of trips with origin home and the income variable from INE, but this time organized in terciles weighted by census section population instead of using the MITMS cuts.

Regarding the computational requirements to run the models: the reproduction models (A, B, C, and D) need nearly 25GB of free RAM if run one by one. In other words: an R Session restarted just before running the chunk with `fepois()` peaks at 25GB of RAM when running that chunk; we recommend making sure you have enough RAM available before running each model.

```{r}
#| message: false
#| warning: false

library(dplyr)
library(tidyr)
library(fixest)
```

## Reproduction models (A, B, C, D)

### Model A: full sample, MITMS income

We build the final panel for model A, this model uses the full sample of trips (we don't need to filter for those we can assign INE income to) and the MITMS income variable. Following Renninger and Cabrera (2026, Supplementary Material, p.7) we turn the UTCI variable into five equal bins from 25 to 50 degrees Celsius, using the 25-30 bin as our reference category.

```{r}
panel_reproduction <- readRDS("../data/model_panels/panel_reproduction.rds")

panel_a <- panel_reproduction |>
  # Group by the original MITMS income:
  group_by(date, id_destination,
           province, # Needed for the cluster
           utci, activity_destination,
           income) |>
  summarise("n_trips" = sum(n_trips), .groups = "drop") |>
  # Create the utci bins:
  mutate(
    "utci_bins" = case_when(
      utci >= 25 & utci < 30 ~ "25-30",
      utci < 35 ~ "30-35",
      utci < 40 ~ "35-40",
      utci < 45 ~ "40-45",
      utci < 50 ~ "45-50")
  ) |>
  select(!utci)

glimpse(panel_a)

# Check if there are any NAs before passing the panel to fepois():
panel_a |> summarise(across(everything(), ~ sum(is.na(.x))))

# There are no NAs
```

While it is not necessary in the computer in which we are running the models, we save the panel as an .rds file to make it easy for smaller computers to run these four models. Most personal computers should be able to run these first four models allocating swap memory (they require nearly 25GB of RAM to run, we recommend making sure you have some more than that). We save the .rds files so that on a smaller personal computer each model can be estimated in a new seesion by restarting R just before running the chunk, which can just load `fixest`, read the panel from the .rds, and run `fepois()` with as much free RAM as possible.

```{r}
# To allow  fepois() to run in a session with as much free RAM as possible,
# we save the panel in an RDS file:
saveRDS(panel_a, "../data/model_panels/panel_a.rds")
```

Originally we used the more basic syntax (with `fml = n_trips ~ utci_bins * activity_destination * income | date + id_destination`), but after working with the results for a while, we realized some things would be easier if we use the `i()` syntax instead, since it “allows for better synergies with other `fixest` functions like `etable` and `coefplot`/`iplot`” ([Bergé and McDermott, 2026, Section 4.2](https://cran.r-project.org/web/packages/fixest/vignettes/fixest_walkthrough.html#4_Interaction_terms)).

```{r}
# If working on a personal computer, restart R before running this chunk.

library(fixest) # Needed for users that restart R before running.

panel_a <- readRDS("../data/model_panels/panel_a.rds")

model_a <- fepois(
  fml = n_trips ~ activity_destination * income +
    i(utci_bins, ref = "25-30"):activity_destination:income | date + id_destination,
  data = panel_a,
  cluster = ~province,
  # Clean intermediary objects as much as possible to free RAM space:
  mem.clean = TRUE,
  nthreads = 48 # Set this value according to your computer
  )

# Create the directory to save the models there:
dir.create("../data/models", showWarnings = FALSE, recursive = TRUE)

# And save it:
saveRDS(model_a, "../data/models/model_a.rds")
```

We read the model to check everything worked:

```{r}
model_a <- readRDS("../data/models/model_a.rds")

etable(model_a)
```

### Model B: subsample, MITMS income

We use the same MITMS income variable as model A, but only the trips that start at home. This is the subsample of trips INE income can be assigned to.

We build the panel for model B, identical to the one for model A except for that restriction. The UTCI bins are the same:

```{r}
# Loading these libraries is needed for users that restart R before running.
library(dplyr)
library(tidyr)

panel_reproduction <- readRDS("../data/model_panels/panel_reproduction.rds")

panel_b <- panel_reproduction |>
  # Keep only the trips that start at home:
  filter(home_trip) |>
  # Group by the original MITMS income:
  group_by(date, id_destination,
           province, # Needed for the cluster
           utci, activity_destination, income) |>
  summarise("n_trips" = sum(n_trips), .groups = "drop") |>
  # Create the utci bins:
  mutate(
    "utci_bins" = case_when(
      utci >= 25 & utci < 30 ~ "25-30",
      utci < 35 ~ "30-35",
      utci < 40 ~ "35-40",
      utci < 45 ~ "40-45",
      utci < 50 ~ "45-50")
  ) |>
  select(!utci)

glimpse(panel_b)

# Check if there are any NAs before passing the panel to fepois():
panel_b |> summarise(across(everything(), ~ sum(is.na(.x))))

# There are no NAs
```

As with model A, we save the panel as an .rds file, so that smaller computers can restart R before estimating:

```{r}
# To allow  fepois() to run in a session with as much free RAM as possible,
# we save the panel in an RDS file:
saveRDS(panel_b, "../data/model_panels/panel_b.rds")
```

```{r}
# If working on a personal computer, restart R before running this chunk.

library(fixest) # Needed for users that restart R before running.

panel_b <- readRDS("../data/model_panels/panel_b.rds")

model_b <- fepois(
  fml = n_trips ~ activity_destination * income +
    i(utci_bins, ref = "25-30"):activity_destination:income | date + id_destination,
  data = panel_b,
  cluster = ~province,
  # Clean intermediary objects as much as possible to free RAM space:
  mem.clean = TRUE,
  nthreads = 48 # Set this value according to your computer
  )

# And save it:
saveRDS(model_b, "../data/models/model_b.rds")
```

We read the model to check everything worked:

```{r}
model_b <- readRDS("../data/models/model_b.rds")

etable(model_b)
```

### Model C: subsample, INE income (MITMS cuts)

Same sample as model B (only the trips that start at home), but the income comes from the INE data of the year the trips were made, using the same brackets as MITMS.

We build the panel for model C, identical to the one for model B except for the income variable:

```{r}
# Loading these libraries is needed for users that restart R before running.
library(dplyr)
library(tidyr)

panel_reproduction <- readRDS("../data/model_panels/panel_reproduction.rds")

panel_c <- panel_reproduction |>
  # Keep only the trips that start at home:
  filter(home_trip) |>
  # Group by the income reconstructed from the INE data:
  group_by(date, id_destination,
           province, # Needed for the cluster
           utci, activity_destination, ine_income_bracket) |>
  summarise("n_trips" = sum(n_trips), .groups = "drop") |>
  # Create the utci bins:
  mutate(
    "utci_bins" = case_when(
      utci >= 25 & utci < 30 ~ "25-30",
      utci < 35 ~ "30-35",
      utci < 40 ~ "35-40",
      utci < 45 ~ "40-45",
      utci < 50 ~ "45-50")
  ) |>
  select(!utci)

glimpse(panel_c)

# Check if there are any NAs before passing the panel to fepois():
panel_c |> summarise(across(everything(), ~ sum(is.na(.x))))

# Check how many trips those NAs are:
panel_c |>
    mutate("missing_ine_income_bracket" = is.na(ine_income_bracket)) |>
    summarise("trips" = sum(n_trips), .by = "missing_ine_income_bracket") |>
    mutate("pct_trips" = trips / sum(trips) * 100)
```

We find NAs in `ine_income_bracket`, as expected, since there were some MITMS zones that didn't include INE section codes. The amount of NAs, however, is residual, they represent just 0.0348% of the trips. We drop them and save the .rds file.

```{r}
# Drop the NAs:
panel_c <- panel_c |> drop_na()

# To allow  fepois() to run in a session with as much free RAM as possible,
# we save the panel in an RDS file:
saveRDS(panel_c, "../data/model_panels/panel_c.rds")
```

```{r}
# If working on a personal computer, restart R before running this chunk.

library(fixest) # Needed for users that restart R before running.

panel_c <- readRDS("../data/model_panels/panel_c.rds")

model_c <- fepois(
  fml = n_trips ~ activity_destination * ine_income_bracket +
    i(utci_bins, ref = "25-30"):activity_destination:ine_income_bracket | date + id_destination,
  data = panel_c,
  cluster = ~province,
  # Clean intermediary objects as much as possible to free RAM space:
  mem.clean = TRUE,
  nthreads = 48 # Set this value according to your computer
  )

# And save it:
saveRDS(model_c, "../data/models/model_c.rds")
```

We read the model to check everything worked:

```{r}
model_c <- readRDS("../data/models/model_c.rds")

etable(model_c)
```

### Model D: subsample, INE terciles

Same sample and same INE income as model C, but cut into population-weighted terciles instead of MITMS's brackets (the cuts made sense with the income distribution of 2020, but not with the other years).

We build the panel for model D, identical to the one for model C except for the income variable:

```{r}
# Loading these libraries is needed for users that restart R before running.
library(dplyr)
library(tidyr)

panel_reproduction <- readRDS("../data/model_panels/panel_reproduction.rds")

panel_d <- panel_reproduction |>
  # Keep only the trips that start at home:
  filter(home_trip) |>
  # Group by the terciles of the INE income:
  group_by(date, id_destination,
           province, # Needed for the cluster
           utci, activity_destination, ine_income_tercile) |>
  summarise("n_trips" = sum(n_trips), .groups = "drop") |>
  # Create the utci bins:
  mutate(
    "utci_bins" = case_when(
      utci >= 25 & utci < 30 ~ "25-30",
      utci < 35 ~ "30-35",
      utci < 40 ~ "35-40",
      utci < 45 ~ "40-45",
      utci < 50 ~ "45-50")
  ) |>
  select(!utci)

glimpse(panel_d)

# Check if there are any NAs before passing the panel to fepois():
panel_d |> summarise(across(everything(), ~ sum(is.na(.x))))

# Check how many trips those NAs are:
panel_d |>
    mutate("missing_ine_income_tercile" = is.na(ine_income_tercile)) |>
    summarise("trips" = sum(n_trips), .by = missing_ine_income_tercile) |>
    mutate("pct_trips" = trips / sum(trips) * 100)
```

Same NAs as in `panel_c`, we just drop them and save the .rds file.

```{r}
panel_d <- panel_d |> drop_na()

# To allow  fepois() to run in a session with as much free RAM as possible,
# we save the panel in an RDS file:
saveRDS(panel_d, "../data/model_panels/panel_d.rds")
```

```{r}
# If working on a personal computer, restart R before running this chunk.

library(fixest) # Needed for users that restart R before running.

panel_d <- readRDS("../data/model_panels/panel_d.rds")

model_d <- fepois(
  fml = n_trips ~ activity_destination * ine_income_tercile +
    i(utci_bins, ref = "25-30"):activity_destination:ine_income_tercile | date + id_destination,
  data = panel_d,
  cluster = ~province,
  # Clean intermediary objects as much as possible to free RAM space:
  mem.clean = TRUE,
  nthreads = 48 # Set this value according to your computer
  )

# And save it:
saveRDS(model_d, "../data/models/model_d.rds")
```

We read the model to check everything worked:

```{r}
model_d <- readRDS("../data/models/model_d.rds")

etable(model_d)
```

### Export reproduction models

We load the four models results and export to LaTeX format using `etable()`, provided by `fixest` for this purpose. We are following [Berge's (2026) vignette on exporting estimation tables](https://cran.r-project.org/web/packages/fixest/vignettes/exporting_tables.html).

```{r}
# If working on a personal computer, restart R before running this chunk.

library(fixest) # Needed for users that restart R before running.

model_a <- readRDS("../data/models/model_a.rds")
model_b <- readRDS("../data/models/model_b.rds")
model_c <- readRDS("../data/models/model_c.rds")
model_d <- readRDS("../data/models/model_d.rds")
```

```{r}
# We set a dictionary renaming the variables used by etable(). We read the
# coefficients with names(coef(model_a)) to check
# what they were, then renamed them in the dictionary:
setFixest_dict(c(
  "n_trips" = "Trips",
  "utci_bins::30-35" = "UTCI 30-35",
  "utci_bins::35-40" = "UTCI 35-40",
  "utci_bins::40-45" = "UTCI 40-45",
  "utci_bins::45-50" = "UTCI 45-50",
  "activity_destinationfrequent_activity" = "Frequent activity",
  "activity_destinationinfrequent_activity" = "Infrequent activity",
  "activity_destinationwork_or_study" = "Work/study",
  "income<10" = "Income <10",
  "income10-15" = "Income 10-15",
  "income>15" = "Income >15",
  "ine_income_bracket<10" = "Income <10",
  "ine_income_bracket10-15" = "Income 10-15",
  "ine_income_bracket>15" = "Income >15",
  "ine_income_tercilelow" = "Low tercile",
  "ine_income_tercilemiddle" = "Middle tercile",
  "ine_income_tercilehigh" = "High tercile",
  "date" = "Date",
  "id_destination" = "District"
  ))
```

```{r}
# Create the directory to save the tables there:
dir.create("../report/tables", showWarnings = FALSE, recursive = TRUE)

# We create one table per temperature bin. Model D has its own tables because
# the income terciles are different categories, so they take extra rows and
# don't let the tables fit in just one page if we put them together with
# models a, b, and c:
for (utci_bin in c("30-35", "35-40", "40-45", "45-50")) {

  # Create the table for each utci bin for models a, b and c:
  etable(model_a,
         model_b,
         model_c,
         # Keep only the coefficients of this bin:
         keep = utci_bin,
         # Name the columns with the model names:
         headers = c("Full sample, MITMS income", "Subsample, MITMS income",
                     "Subsample, INE income"),
         # Print only the statistics we want. By default etable() also prints
         # the BIC, which we don't need. On top of that, models B, C and F
         # printed a BIC below 1 while the others had BICS of billions, computing
         # it with BIC() gave numbers in the same orders of magnitude for all
         # of them, so it must be a problem with `fixest`:
         fitstat = ~ n + sq.cor + pr2,
         title = paste0("Change in trips at UTCI ",
                        utci_bin,
                        " °C, against the 25-30 °C reference. ",
                        "Sample: trips terminating in the district"),
         # The LaTeX label, we will use this to reference the table. The same
         # models are estimated on the two samples, so the label says which one
         # this table comes from:
         label = paste0("table_models_abc_terminating_", utci_bin),
         fontsize = "small",
         # The model names make the table wider than the text block, we use
         # adjustbox to allow LaTeX to adjust the tables, the package
         # `adjustbox` must be loaded in LaTeX:
         adjustbox = TRUE,
         # To make the etable in LaTeX format:
         tex = TRUE,
         file = paste0("../report/tables/table_models_abc_", utci_bin, ".tex"),
         # Overwrite the file instead of appending to it:
         replace = TRUE)

  # Create the table for each utci bin for model D:
  etable(model_d,
         keep = utci_bin,
         headers = "Subsample, INE terciles",
         # Same as above, we print only the statistics we want (no BIC):
         fitstat = ~ n + sq.cor + pr2,
         title = paste0("Change in trips at UTCI ", utci_bin,
                        " °C by income tercile, against the 25-30 °C reference"),
         label = paste0("table_model_d_", utci_bin),
         fontsize = "small",
         tex = TRUE,
         file = paste0("../report/tables/table_model_d_",
                       utci_bin, ".tex"),
         replace = TRUE)
}
```

These models perform worse than the models using the within district trips panel, results from the models in script `7b_estimate_the_models_within_trips.qmd` are closer to Renninger and Cabrera (2026), which provides further evidence that the within district trips panel was the panel they used.

