---
title: "Plotting the results"
author: "Miguel Agenjo Montejo"
format: html
editor: visual
---

## Overview and set up

The models are estimated and saved in `07b_estimate_the_models_within_district.qmd`. Here we only read them, pull their coefficients into a small tibble, and generate and save the plots. For easier reproducibility in smaller personal computers, we never load more than one model in memory at a time (they weight between 1GB and 14.5 GB).

```{r}
#| message: false
#| warning: false

library(fixest)
library(broom)
library(dplyr)
library(tidyr)
library(stringr)
library(tibble)
library(ggplot2)
library(RColorBrewer)
library(knitr)


# We defined a custom theme for all the figures in the masters thesis:
source("../dataviz_theme/theme_tfm.R")

# Create the directory to save the figures there:
dir.create("../report/figures", showWarnings = FALSE, recursive = TRUE)
```

## Labels and colors

We set readable labels and model names, so that the plots do not show the raw names:

```{r}
# More easily readable model names:
model_labels <- c("model_a" = "A: full sample",
                  "model_b" = "B: subsample",
                  "model_c" = "C: INE income",
                  "model_d" = "D: INE terciles")

# Activity labels:
activity_labels <- c("work_or_study" = "Work/study",
                     "frequent_activity" = "Frequent",
                     "infrequent_activity" = "Infrequent")


# Age labels:
age_labels <- c("0-25" = "Age 0-25",
                "25-45" = "Age 25-45",
                "45-65" = "Age 45-65",
                "65-100" = "Age 65-100")
```

And some of the colors from `RColorBrewer` which we use manually in some of the plots (see why we chose these colors in `../dataviz_theme/dataviz_decisions.qmd`).

Colors for the income categories:

```{r}
# Colors for the income categories (print hex codes):
brewer.pal(4, "YlGnBu")
# Save them:
income_colours <- c("Low" = "#A1DAB4",
                    "Middle" = "#41B6C4",
                    "High" = "#225EA8")

# Colors for the UTCI bins (print hex codes):
brewer.pal(9, "YlOrRd")[c(4, 6, 7, 9)]
# Save them:
utci_colours <- c("30-35" = "#FEB24C",
                  "35-40" = "#FC4E2A",
                  "40-45" = "#E31A1C",
                  "45-50" = "#800026")

# Colors for the age groups (print hex codes):
brewer.pal(8, "Purples")[c(4, 6, 8)]
# Save them:
age_colours <- c("25-45" = "#BCBDDC",
                 "45-65" = "#807DBA",
                 "65-100" = "#4A1486")
```

## Reproduction models plots

### Load data

We read the coefficients with `tidy()`, which returns them in a tibble, with their names in a column and their confidence intervals already computed from the clustered standard errors of the model. We generate a tibble with all the coefficients for models A to D using a for loop that appends the rows of each model one at a time, removing the previous model before laoding the next one to avoid overloading the session memory in smaller personal computers:

```{r}
# A vector with all the models we read:
models_abcd <- c("model_a", "model_b", "model_c", "model_d")

# Create the emty tibble with the coefficients:
coefficients_abcd_raw <- tibble()

# For every model:
for (one_model in models_abcd) {

  # Load the .rds file:
  model <- readRDS(paste0("../data/models/", one_model, "_within.rds"))

  # Extract the coefficients with their confidence intervals:
  coefficients_abcd_raw <- tidy(model, conf.int = TRUE) |>
    # Keep only the coefficients with UTCI interactions:
    filter(str_detect(term, "utci_bins::")) |>
    # Create a column indicating what model the coefficients come from:
    mutate("model" = one_model) |>
    # And add the rows to the already existing tibble:
    bind_rows(coefficients_abcd_raw)

  # Free the memory before reading the next model:
  rm(model)
}

glimpse(coefficients_abcd_raw)
```

As we can see, the terms have three parts: activity destination, income, and UTCI bin, we split them so that each of them has their own column, and clean their values so that they look better in the plots:

```{r}
coefficients_abcd <- coefficients_abcd_raw |>
  # The three parts are joined by ":" so we separate them:
  separate_wider_delim(
    term, delim = ":",
    # We include two NAs for the "utci_bins" text and the column that results
    # of splitting  the "::", this way those columns are not created:
    names = c("activity_destination", "income", NA, NA, "utci_bins")) |>
  # The values still have the name of the original variable in front of
  # them, we remove those:
  mutate(
    "activity_destination" = str_remove(activity_destination,
                                        "activity_destination"),
    # The income variable has a different name in each model:
    "income" = str_remove(income,
                          "^ine_income_tercile|^ine_income_bracket|^income"),
    # Brackets (A-C) and terciles (D) are different variables, but they order
    # the zones the same way, so we translate them into a common scale:
    "income_level" = case_when(
      income %in% c("<10", "low") ~ "Low",
      income %in% c("10-15", "middle") ~ "Middle",
      income %in% c(">15", "high") ~ "High") |>
      factor(levels = c("Low", "Middle", "High")))

glimpse(coefficients_abcd)
```

Since the models are Poisson, the estimates are changes in the log count, to make them easier to interpret, we exponentiate them to convert them into incidence rate ratios. Each of them is a ratio against the reference bin (25-30 °C) within the same activity and income level:

```{r}
coefficients_abcd <- coefficients_abcd |>
  mutate("irr" = exp(estimate),
         "irr_low" = exp(conf.low),
         "irr_high" = exp(conf.high))

glimpse(coefficients_abcd)
```

We want to visually distinguish coeffients that are statistically distinguishable from no effect (the confidence interval doesn't include and IRR of 1) by giving more visual weight to their points. We will do that by leaving them filled (leaving hollow those points whose confidence interval crosses an IRR 1). In order to know whatpoints to fill and which to leave blank we create a binary variable with that information:

```{r}
data_plots_reproduction <- coefficients_abcd |>
  mutate("distinguishable" = if_else(irr_low > 1 | irr_high < 1,
                                     "Yes",
                                     "No"))
```

### Plot 1a

```{r}
data_plot_1a <- data_plots_reproduction |>
  filter(model == "model_a")
```

```{r}
#| fig-width: 5.9
#| fig-height: 2.6

# Create the plot:
plot_1a <- ggplot(data_plot_1a,
                 aes(y = utci_bins, x = irr,
                     colour = income_level,
                     # Group by income only, if we don't explictly do this, 
                     # the points are also grouped by shape and they no 
                     # longer sit on top of their own CI line:
                     group = income_level)) +

  # A dashed line for IRR value of 1:
  geom_vline(xintercept = 1, linetype = "dashed", colour = "grey50") +

  # Add the lines for the confidence interval ranges:
  geom_linerange(aes(xmin = irr_low, xmax = irr_high),
                 linewidth = 0.3,
                 position = position_dodge(width = 0.5)) +
  
  # Add the points in the point estimates, all of them filled:
  geom_point(aes(fill = income_level),
             shape = 21, size = 1.6, stroke = 0.5,
             position = position_dodge(width = 0.5)) +
  # And add another point on top, this time filled in white. We will make
  # the size = 0 when their CI crosses IRR of 1, so that we see the filled
  # point or a "hollow" (white) point depending on the case:
  geom_point(aes(size = distinguishable),
             shape = 21, fill = "white", stroke = 0.5,
             position = position_dodge(width = 0.5)) +

  # Create the grid with one model per row and an activity per column:
  facet_grid(model ~ activity_destination,
             labeller = labeller(activity_destination = activity_labels,
                                 model = model_labels),
             # Place the name of the models to the left:
             switch = "y") +
  # And place the utci bins to the right:
  scale_y_discrete(position = "right") +
  
  # Use a logaritmic scale so that the visual representation shows the
  # effects symmetrically in both directions:
  scale_x_log10(breaks = c(0.25, 0.5, 1, 2, 4)) +

  # Set the color of the interval lines and the points contours:
  scale_colour_manual(values = income_colours) +
  # Set the fill of the points:
  scale_fill_manual(values = income_colours) +
  # When the CI crosses an IRR of 1, we make the size 0 so that the hollow
  # point does not appear, leaving visible the filled 1:
  scale_size_manual(values = c("Yes" = 0, "No" = 1.6)) +

  # Edit the legend so that we get just one legend that reflects both the
  # fill and the color:
  guides(fill = "none", size = "none",
         colour = guide_legend(override.aes = list(shape = 21,
                                                   fill = income_colours))) +

  # Define the labels:
  labs(y = "UTCI bin (°C)",
       x = "Change in trips relative to the 25-30 °C bin (IRR, log scale)",
       colour = "Income:",
       caption = "Hollow point: confidence interval crosses IRR = 1") +

  # Our custom theme:
  theme_tfm +
  # And some tweaks:
  theme(
    # Remove the horizontal grid lines:
    panel.grid.major.y = element_blank(),
    # Put the strip outside of the facets:
    strip.placement = "outside",
    # Place the legend at the bottom:
    legend.position = "bottom",
    # Move the legend up a little bit:
    legend.box.spacing = unit(0, "pt"),
    # And the caption right under it, centered instead of to the right:
    plot.caption = element_text(hjust = 0.5, margin = margin(t = -6)))

plot_1a

# Save it:
ggsave_tfm("../report/figures/plot1a_irr_by_income_pointrange_within.pdf",
       plot_1a,
       width = 5.9, height = 2.6,
       create.dir = TRUE)
```

### Plot 1b

```{r}
data_plot_1b <- data_plots_reproduction |>
  filter(model != "model_a")
```

```{r}
#| fig-width: 5.9
#| fig-height: 4.5

# Create the plot:
plot_1b <- ggplot(data_plot_1b,
                 aes(y = utci_bins, x = irr,
                     colour = income_level,
                     # Group by income only, if we don't explictly do this, 
                     # the points are also grouped by shape and they no 
                     # longer sit on top of their own CI line:
                     group = income_level)) +

  # A dashed line for IRR value of 1:
  geom_vline(xintercept = 1, linetype = "dashed", colour = "grey50") +

  # Add the lines for the confidence interval ranges:
  geom_linerange(aes(xmin = irr_low, xmax = irr_high),
                 linewidth = 0.3,
                 position = position_dodge(width = 0.5)) +
  
  # Add the points in the point estimates, all of them filled:
  geom_point(aes(fill = income_level),
             shape = 21, size = 1.6, stroke = 0.5,
             position = position_dodge(width = 0.5)) +
  # And add another point on top, this time filled in white. We will make
  # the size = 0 when their CI crosses IRR of 1, so that we see the filled
  # point or a "hollow" (white) point depending on the case:
  geom_point(aes(size = distinguishable),
             shape = 21, fill = "white", stroke = 0.5,
             position = position_dodge(width = 0.5)) +

  # Create the grid with one model per row and an activity per column:
  facet_grid(model ~ activity_destination,
             labeller = labeller(activity_destination = activity_labels,
                                 model = model_labels),
             # Place the name of the models to the left:
             switch = "y") +
  # And place the utci bins to the right:
  scale_y_discrete(position = "right") +
  
  # Use a logaritmic scale so that the visual representation shows the
  # effects symmetrically in both directions. These three models sit much
  # closer to 1 than model A, so they need their own breaks:
  scale_x_log10(breaks = c(0.5, 0.75, 1, 1.25)) +

  # Set the color of the interval lines and the points contours:
  scale_colour_manual(values = income_colours) +
  # Set the fill of the points:
  scale_fill_manual(values = income_colours) +
  # When the CI crosses an IRR of 1, we make the size 0 so that the hollow
  # point does not appear, leaving visible the filled 1:
  scale_size_manual(values = c("Yes" = 0, "No" = 1.6)) +

  # Edit the legend so that we get just one legend that reflects both the
  # fill and the color:
  guides(fill = "none", size = "none",
         colour = guide_legend(override.aes = list(shape = 21,
                                                   fill = income_colours))) +

  # Define the labels:
  labs(y = "UTCI bin (°C)",
       x = "Change in trips relative to the 25-30 °C bin (IRR, log scale)",
       colour = "Income:",
       caption = "Hollow point: confidence interval crosses IRR = 1") +

  # Our custom theme:
  theme_tfm +
  # And some tweaks:
  theme(
    # Remove the horizontal grid lines:
    panel.grid.major.y = element_blank(),
    # Put the strip outside of the facets:
    strip.placement = "outside",
    # Place the legend at the bottom:
    legend.position = "bottom",
    # Move the legend up a little bit:
    legend.box.spacing = unit(0, "pt"),
    # And the caption right under it, centered instead of to the right:
    plot.caption = element_text(hjust = 0.5, margin = margin(t = -6)))

plot_1b

# Save it:
ggsave_tfm("../report/figures/plot1b_irr_by_income_pointrange_within.pdf",
       plot_1b,
       width = 5.9, height = 4.5,
       create.dir = TRUE)
```

## Intersectionality models plots

### Load data

```{r}
# Load the model (this object is 16.2GB):
model_full <- readRDS("../data/models/model_full_within.rds")

# Extract the coefficients with their confidence intervals:
coefficients_full_raw <- tidy(model_full, conf.int = TRUE) |>
  # Keep only the coefficients with UTCI interactions:
  filter(str_detect(term, "utci_bins::"))

# Free the memory, the model is a big object:
rm(model_full)

glimpse(coefficients_full_raw)
```

The terms of this model have five parts: activity at destination, income tercile, age group, sex, and UTCI bin. We split them and clean their values the same way we did for models A to D:

```{r}
coefficients_full <- coefficients_full_raw |>
  # The five parts are joined by ":" so we separate them:
  separate_wider_delim(
    term, delim = ":",
    # We include two NAs for the "utci_bins" text and the column that results
    # of splitting the "::", this way those columns are not created:
    names = c("activity_destination", "income", "age", "sex", NA, NA,
              "utci_bins")) |>
  
  # The values still have the name of the original variable in front of
  # them, we remove those:
  mutate(
    "activity_destination" = str_remove(activity_destination,
                                        "^activity_destination"),
    "income" = str_remove(income, "^ine_income_tercile"),
    "age" = str_remove(age, "^age"),
    "sex" = str_remove(sex, "^sex"),
    # The same scale we used for models A to D:
    "income_level" = case_when(
      income == "low" ~ "Low",
      income == "middle" ~ "Middle",
      income == "high" ~ "High") |>
      factor(levels = c("Low", "Middle", "High")),
    
    # The activities in the order we always use them, so that the facets do not
    # come out alphabetically:
    "activity_destination" = factor(
      activity_destination,
      levels = c("work_or_study", "frequent_activity",
                 "infrequent_activity")),
    
    # And the same conversion into incidence rate ratios:
    "irr" = exp(estimate),
    "irr_low" = exp(conf.low),
    "irr_high" = exp(conf.high)
    )

glimpse(coefficients_full)
```

### Plot 2

This is the same plot as 1, but this time we plot gender (using two different kinds of points) and age (using facets) too:

```{r}
# Create the same binary variable we built for plots 1a and 1b:
data_plot_2 <- coefficients_full |>
  mutate("distinguishable" = if_else(irr_low > 1 | irr_high < 1, "Yes", "No"))
```

```{r}
#| fig-width: 5.9
#| fig-height: 5.5

# Create the plot:
plot_2 <- ggplot(data_plot_2,
                 aes(y = utci_bins, x = irr, colour = income_level,
                     # Group by income and gender, otherwise the points are
                     # also grouped by shape and they no longer sit on top of
                     # their own CI line:
                     group = interaction(income_level, sex))) +

  # A dashed line for IRR value of 1:
  geom_vline(xintercept = 1, linetype = "dashed", colour = "grey50") +

  # Add the lines for the confidence interval ranges:
  geom_linerange(aes(xmin = irr_low, xmax = irr_high), linewidth = 0.22,
                 position = position_dodge(width = 0.9)) +

  # As in plots 1a and 1b, all the points filled, and a white point one on top
  # of the one in which the CI crosses 1. This time we also change the shape 
  # based on the gender, circle (21) for women, triangle (24) for men:
  geom_point(aes(shape = sex, fill = income_level),
             size = 1.5, stroke = 0.5,
             position = position_dodge(width = 0.9)) +
  geom_point(aes(shape = sex, size = distinguishable),
             fill = "white", stroke = 0.5,
             position = position_dodge(width = 0.9)) +

  # Crete a grid with activities as rows and age as columns:
  facet_grid(activity_destination ~ age,
             labeller = labeller(activity_destination = activity_labels,
                                 age = age_labels),
             # Same as above, so that the strips appear outside:
             switch = "y") +

  # Move the bin labels to the right and add some extra space so that the
  # points at the top and bottom of each facet don't sit too close to the edge:
  scale_y_discrete(position = "right", expand = expansion(add = 0.7)) +
  # Use a logaritmic scale so that the visual representation shows the
  # effects symmetrically in both directions:
  scale_x_log10(breaks = c(0.75, 1, 1.25)) +

  # Set the colors of the contour of the points and the interval lines:
  scale_colour_manual(values = income_colours) +
  # Set the shape assigned to each value (circle for women, triangle for men):
  scale_shape_manual(values = c("female" = 21, "male" = 24),
                     labels = c("female" = "Women", "male" = "Men")) +
  # Set the color of the fill for the points:
  scale_fill_manual(values = income_colours) +

  # Same as in plots 1a and 1b, we make the size of the point on top 0 when the
  # CI crosses 1 to get the effect we want (filled when it doesn't, white when
  # it does):
  scale_size_manual(values = c("Yes" = 0, "No" = 1.5)) +

  # Edit the legend so that we get just one legend that explains the fill,
  # the color, and the shape:
  guides(
    fill = "none", size = "none",
    colour = guide_legend(order = 1,
                          override.aes = list(shape = 21,
                                              fill = income_colours)),
    shape = guide_legend(order = 2,
                         override.aes = list(fill = "black",
                                             colour = "black"))) +

  # Define the labels:
  labs(y = "UTCI bin (°C)",
       x = "Change in trips relative to the 25-30 °C bin (IRR, log scale)",
       colour = "Income:", shape = "Sex:",
       caption = "Hollow point: confidence interval crosses IRR = 1") +

  # Custom theme:
  theme_tfm +
  # And some tweaks:
  theme(panel.grid.major.y = element_blank(),
        strip.placement = "outside",
        legend.position = "bottom",
        # Move the legend up a little bit:
        legend.box.spacing = unit(0, "pt"),
        # And the caption right under it, centered instead of to the right:
        plot.caption = element_text(hjust = 0.5, margin = margin(t = -6)))

plot_2

# Save it:
ggsave_tfm("../report/figures/plot2_irr_by_age_sex_pointrange_within.pdf", plot_2,
       width = 5.9, height = 5.5)
```

### Plot 3

This plot intends to help visualize the difference (lack of) between men and women:

```{r}
data_plot_3 <- coefficients_full |>
  # Select the relevant variables:
  select(activity_destination, income_level, age, sex, utci_bins, irr) |>
  # We don't trust the data on the 0-25 group, so we drop it:
  filter(age != "0-25") |>
  # Pivot wider so that we get one row per cell, 
  # with the two sexes side by side in their own columns:
  pivot_wider(names_from = sex, values_from = irr)
```

```{r}
#| fig-width: 5.9
#| fig-height: 6

# Create the plot:
plot_3 <- ggplot(data_plot_3, aes(x = male, y = female,
                                  colour = utci_bins, shape = age)) +

  # Draw the diagonal line (where the points should fall if men and women had
  # the same values, the distance from it is the distance between sexes:
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey50") +
  
  ## Add the points:
  geom_point(size = 1.6) +

  # One row per income tercile, and one column per activity:
  facet_grid(income_level ~ activity_destination,
             labeller = labeller(activity_destination = activity_labels,
                                 income_level = c("Low" = "Low income",
                                                  "Middle" = "Middle income",
                                                  "High" = "High income"))) +

  # Set the color of the points based on the UTCI bin:
  scale_colour_manual(values = utci_colours) +
  
  # We use a different shapes this time, so that the reader quickly understands
  # shape does not mean gender this time:
  scale_shape_manual(values = c("25-45" = 3, "45-65" = 4, "65-100" = 8),
                     labels = age_labels) +

  # Both axes are the same quantity, so they need the same scale: without
  # this the diagonal stops being the 45º:
  coord_fixed(ratio = 1) +
  labs(x = "Men (IRR)", y = "Women (IRR)",
       colour = "UTCI bin (°C):", shape = "Age:") +

  # Side by side the two legends are wider than the plot and the last UTCI bin
  # is cut off, so we put age first and the bins underneath:
  guides(shape = guide_legend(order = 1),
         colour = guide_legend(order = 2)) +

  # Custom theme:
  theme_tfm +
  # And some tweaks:
  theme(
        # Add some space between panels, so that the labels don't overlap:
        panel.spacing.x = unit(1, "lines"),
        legend.position = "bottom",
        # One legend per row instead of both on the same one so that they fit
        # the page length:
        legend.box = "vertical",
        # Get the two lines in the legend closer.
        legend.spacing.y = unit(0, "pt"),
        # Move the legend up a little bit:
        legend.box.spacing = unit(2, "pt"),
        legend.margin = margin(0, 0, 0, 0))

plot_3

# Save it:
ggsave_tfm("../report/figures/plot3_sex_diagonal_scatter_within.pdf", plot_3,
       width = 5.9, height = 6)
```

### Plot 4

This plot intends to help visualize the difference between age groups:

```{r}
data_plot_4 <- coefficients_full |>
  # Select the relevant variables:
  select(activity_destination, income_level, age, sex, utci_bins, estimate) |>
  # We don't trust the data on the 0-25 group, so we drop it:
  filter(age != "0-25") |>
  # We found no relevant gender differences, so we average them to plot just
  # one line (not one for each gender):
  summarise("estimate" = mean(estimate),
            .by = c(activity_destination, income_level, age, utci_bins)) |>
  # Pivot wirder to get one row per cell, with the three terciles in their
  # own columns:
  pivot_wider(names_from = income_level, values_from = estimate) |>
  # The distance between the two ends of the income distribution:
  mutate("gap" = Low - High)
```

```{r}
#| fig-width: 5.9
#| fig-height: 3

# Create the plot:
plot_4 <- ggplot(data_plot_4, aes(x = utci_bins, y = gap,
                                  colour = age, group = age)) +

  # A dashed line at 0 (reprsenting where both the low and high income tercile   
  # respond to heat in the same way:
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +

  # We want to see the slope (how the distance between age groups grows
  # as UTCI rise, so we plot a line too):
  geom_line(linewidth = 0.5) +
  geom_point(size = 1.6) +

  # One panel per activity (we don't have gender, so don't need a grid):
  facet_wrap(~ activity_destination, nrow = 1,
             labeller = labeller(activity_destination = activity_labels)) +

  # Set the color for age groups:
  scale_colour_manual(values = age_colours,
                      labels = age_labels) +

  # Define the labels:
  labs(x = "UTCI bin (°C)",
       y = "Gap between the lowest and\nthe highest tercile (log points)",
       colour = "Age:") +

  # Custom theme:
  theme_tfm +
  # And some tweaks:
  theme(legend.position = "bottom",
        # Put the legend a bit closer:
        legend.box.spacing = unit(1, "pt"),
        # Make the left label text a little bit smaller:
        axis.title.y = element_text(size = 9)
        )

plot_4

# Save it:
ggsave_tfm("../report/figures/plot4_income_gap_by_age_lines_within.pdf", plot_4,
       width = 5.9, height = 3)
```

## Counting the significant cells

The figures above show which cells can be told apart from no effect, but not how many there are. We count them here, since it is a number we report in the text when commenting on the plots:

```{r}
coefficients_full |>
  summarise("crossings" = n(),
            "significant_001" = sum(p.value <= 0.01),
            "significant_005" = sum(p.value <= 0.05),
            "significant_01" = sum(p.value <= 0.1))
```

Out of 288 crossings, 63 are significant at the 0.01 level, 99 at the 0.05 level, and 120 at the 0.1 level (each count includes the previous one).

