---
title: "Descriptive table and heat days"
author: "Miguel Agenjo Montejo"
format: html
editor: visual
---

## Overview and set up

In this document we build the equivalent to table S1 in the Supplementary Materials of Renninger and Cabrera (2026), and a plot to visualize the number of district-days that fall in each UTCI bin.

```{r}
#| message: false
#| warning: false

library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(DBI)
library(duckdb)
library(knitr)
library(kableExtra)

# We defined a custom theme for all the figures in the masters thesis:
source("../dataviz_theme/theme_tfm.R")
```

## Descriptive statistics (trips terminating in the district)

We build a table equivalent to table S1 in the Supplementary Material of Renninger and Cabrera (2026), so that both can be read side by side.

The panels the models use only keep the district-days between 25 and 50 degrees, so the counts here come from the intermediate table `06_build_panels_for_models.qmd` writes, which still holds every district-day of the period.

```{r}
#| message: false

model_panel_con <- dbConnect(
  duckdb(),
  dbdir = "../data/model_panels/model_panel.duckdb",
  read_only = TRUE # We won't edit the database in this script
  )

# We set the maximum CPU threads and RAM usage, we are working on a powerful
# computer, these values should be adapted to your machine:
dbExecute(model_panel_con, "SET memory_limit = '128GB'")
dbExecute(model_panel_con, "SET threads = 24")

# Load the lazy table:
trips_panel_demographic <- tbl(model_panel_con,
                               "trips_panel_demographic")
```

Each of their three panels in Renninger and Cabrera (2026) adds up to 100% on its own, and the totals differ, meaning they drop the NA values and compute the shares over the trips that are left. We do the same.

Define a funciton to the same operation to the three strata:

```{r}
descriptives_by_stratum <- function(stratum) {

  # The same subset for the three queries below: the districts outside Spain
  # have no UTCI, NAs are dropped. The stratum is embraced so that the
  # function looks inside it for the column name (see R4DS 25.3 here:
  # https://r4ds.hadley.nz/functions.html#data-frame-functions), and renamed
  # so that the three variables can be put together later with bind_rows():
  stratum_panel <- trips_panel_demographic |>
    filter(!is.na(utci), !is.na({{stratum}})) |>
    rename("stratum" = {{stratum}})

  # Sum the trips by stratum and activity at destination:
  stratum_trips <- stratum_panel |>
    group_by(stratum, activity_destination) |>
    summarise("trips" = sum(n_trips), .groups = "drop") |>
    collect()

  # Observations in our sample are district-day cells, so we count the distinct
  # date and district each group appears in:
  stratum_observations <- stratum_panel |>
    distinct(stratum, activity_destination, date, id_destination) |>
    count(stratum, activity_destination, name = "observations") |>
    collect()

  # And count distinct districts, the last column in their table S1:
  stratum_districts <- stratum_panel |>
    distinct(stratum, activity_destination, id_destination) |>
    count(stratum, activity_destination, name = "districts") |>
    collect()

  # Put the three together:
  stratum_trips |>
    left_join(stratum_observations,
              by = join_by(stratum, activity_destination)) |>
    left_join(stratum_districts,
              by = join_by(stratum, activity_destination)) |>
    mutate(
      # Add the trip share (calculated inside each stratum)
      "share" = 100 * trips / sum(trips),
      # And the stratum name:
      "stratum" = as.character(stratum))
}
```

Apply the function to the three strata and bind them together in a single table:

```{r}
age_group <- descriptives_by_stratum(age) |>
  mutate("panel" = "Age group")

income_bracket <- descriptives_by_stratum(income) |>
  mutate("panel" = "Income")

sex_group <- descriptives_by_stratum(sex) |>
  mutate("panel" = "Sex")

table_descriptives <- bind_rows(age_group, 
                                income_bracket,
                                sex_group) |>
  # Their units: billions of trips and millions of observations:
  mutate("trips" = trips / 1e9,
         "observations" = observations / 1e6) |>
  # Relocate the columns so that they follow the same order as theirs:
  relocate(panel) |>
  relocate(share, .after = trips)

table_descriptives

dbDisconnect(model_panel_con, shutdown = TRUE)
```

And we export it to LaTeX, so that it can be read next to theirs:

```{r}
#| message: false

# Define readable names for the activities, as in 08_plot_the_models.qmd:
activity_labels <- c("work_or_study" = "Work or study",
                     "frequent_activity" = "Frequent",
                     "infrequent_activity" = "Infrequent")

table_descriptives_latex <- table_descriptives |>
  # as.character() because DuckDB returns the column as a factor, and
  # indexing a named vector with a factor uses the level number and not
  # the text, which left the work or study rows without a label:
  mutate("activity_destination" =
           activity_labels[as.character(activity_destination)]) |>
  # Write the name of the stratum and of the panel only on the first of their
  # rows, so that the table reads as the three blocks of theirs:
  mutate("stratum" = if_else(row_number() == 1, stratum, ""),
         .by = c(panel, stratum)) |>
  mutate("panel" = if_else(row_number() == 1, panel, ""),
         .by = panel)

# A line between the three panels of the table. kable writes whatever linesep
# says after each row, taking the values in order, so we ask for a line on the
# last row of each panel: the one whose next row already belongs to another one.
# "\midrule" is the horizontal rule of the booktabs LaTeX package:
block_rules <- if_else(lead(table_descriptives_latex$panel, default = "") != "",
                       "\\midrule", "")

# And generate the table with kable():
table_descriptives_latex |>
  kable(format = "latex", booktabs = TRUE,
        # Define column names:
        col.names = c("", "", "Activity", "Total (billions)", "Share (%)",
                      "Observations (millions)", "Districts"),
        # Define the number of decimals we want in each column:
        digits = c(0, 0, 0, 2, 1, 2, 0),
        # This replaces the blank line kable adds every five rows:
        linesep = block_rules) |>
  # We use kableExtra's function add_header_above() to add the "trips" header
  # above the two columns referring to trips:
  add_header_above(c(" " = 3, "Trips" = 2, " " = 2)) |>
  # Give it no caption and no label, because with them kable wraps the table in
  # a block that never prints inside the adjustbox the report needs to fit its
  # width. We write the caption, the label and the adjustbox by hand in the
  # final document.
  as.character() |>
  # Set the file path:
  writeLines("../report/tables/table_descriptives_terminating.tex")
```

## Descriptive statistics (within trips)

Just for comparison (not exported to the final document) we build the same table for just trips within district.

```{r}
#| message: false

model_panel_within_con <- dbConnect(
  duckdb(),
  dbdir = "../data/model_panels/model_panel_within.duckdb",
  read_only = TRUE # We won't edit the database in this script
  )

# We set the maximum CPU threads and RAM usage, we are working on a powerful
# computer, these values should be adapted to your machine:
dbExecute(model_panel_within_con, "SET memory_limit = '128GB'")
dbExecute(model_panel_within_con, "SET threads = 24")

# Load the lazy table:
trips_panel_demographic <- tbl(model_panel_within_con,
                               "trips_panel_demographic")
```

Each of their three panels in Renninger and Cabrera (2026) adds up to 100% on its own, and the totals differ, meaning they drop the NA values and compute the shares over the trips that are left. We do the same.

Define a funciton to the same operation to the three strata:

```{r}
descriptives_by_stratum <- function(stratum) {

  # The same subset for the three queries below: the districts outside Spain
  # have no UTCI, NAs are dropped. The stratum is embraced so that the
  # function looks inside it for the column name (see R4DS 25.3 here:
  # https://r4ds.hadley.nz/functions.html#data-frame-functions), and renamed
  # so that the three variables can be put together later with bind_rows():
  stratum_panel <- trips_panel_demographic |>
    filter(!is.na(utci), !is.na({{stratum}})) |>
    rename("stratum" = {{stratum}})

  # Sum the trips by stratum and activity at destination:
  stratum_trips <- stratum_panel |>
    group_by(stratum, activity_destination) |>
    summarise("trips" = sum(n_trips), .groups = "drop") |>
    collect()

  # Observations in our sample are district-day cells, so we count the distinct
  # date and district each group appears in:
  stratum_observations <- stratum_panel |>
    distinct(stratum, activity_destination, date, id_destination) |>
    count(stratum, activity_destination, name = "observations") |>
    collect()

  # And count distinct districts, the last column in their table S1:
  stratum_districts <- stratum_panel |>
    distinct(stratum, activity_destination, id_destination) |>
    count(stratum, activity_destination, name = "districts") |>
    collect()

  # Put the three together:
  stratum_trips |>
    left_join(stratum_observations,
              by = join_by(stratum, activity_destination)) |>
    left_join(stratum_districts,
              by = join_by(stratum, activity_destination)) |>
    mutate(
      # Add the trip share (calculated inside each stratum)
      "share" = 100 * trips / sum(trips),
      # And the stratum name:
      "stratum" = as.character(stratum))
}
```

Apply the function to the three strata and bind them together in a single table:

```{r}
age_group <- descriptives_by_stratum(age) |>
  mutate("panel" = "Age group")

income_bracket <- descriptives_by_stratum(income) |>
  mutate("panel" = "Income")

sex_group <- descriptives_by_stratum(sex) |>
  mutate("panel" = "Sex")

table_descriptives <- bind_rows(age_group, 
                                income_bracket,
                                sex_group) |>
  # Their units: billions of trips and millions of observations:
  mutate("trips" = trips / 1e9,
         "observations" = observations / 1e6) |>
  # Relocate the columns so that they follow the same order as theirs:
  relocate(panel) |>
  relocate(share, .after = trips)

table_descriptives

dbDisconnect(model_panel_within_con, shutdown = TRUE)
```

## Heat days by district

```{r}
# The third colour of Set2, the blue we use when a single colour has to fill
# large areas:
single_colour <- "#8DA0CB"

# The panel of the reproduction models, already restricted to 25-50 degrees:
panel_reproduction_within <- readRDS(
  "../data/model_panels/panel_reproduction_within.rds")

heat_day_district <- panel_reproduction_within |>
  # This panel keeps the raw UTCI, the bins are created in
  # 07b_estimate_the_models_within_district.qmd, so we repeat them here:
  mutate(
    "utci_bins" = case_when(
      utci >= 25 & utci < 30 ~ "25-30",
      utci < 35 ~ "30-35",
      utci < 40 ~ "35-40",
      utci < 45 ~ "40-45",
      utci < 50 ~ "45-50")
  ) |>
  # Count distinct district-days by utci bin:
  distinct(date, id_destination, utci_bins) |>
  count(utci_bins, name = "district_days")

heat_day_district
```

```{r}
#| fig-width: 5.9
#| fig-height: 4.4

# And plot it:
plot_heat_exposure <- ggplot(heat_day_district,
                             aes(y = utci_bins, x = district_days)) +
  # Bars fill large areas, so we use the blue of Set2 and not black:
  geom_col(fill = single_colour, width = 0.6) +
  scale_x_continuous(
    # One line every 100,000 district-days:
    breaks = scales::breaks_width(100000),
    # Short scale labels, so that 300000 prints as 300K and not as 3e+05:
    labels = scales::label_number(scale_cut = scales::cut_short_scale())) +
  # Add the labels:
  labs(y = "UTCI bin (°C)",
       x = "District-days") +
  theme_tfm +
  theme(# The horizontal grid lines add nothing here:
        panel.grid.major.y = element_blank(),
        # The vertical ones are the breaks we set above:
        panel.grid.minor.x = element_blank())

plot_heat_exposure

# Save it:
ggsave_tfm("../report/figures/day_districts_heat_exposure_barplot.pdf",
       plot_heat_exposure,
       width = 5.9, height = 4.4)
```
