---
title: "Building the panel to reproduce table S2"
author: "Miguel Agenjo Montejo"
format: html
editor: visual
---

## Overview and set up

In this document we build the panel to try to reproduce table S2 in the Supplementary Material of Renninger and Cabrera (2026), which estimates the effect of a day above 40 degrees on the flows within a district and between districts.

```{r}
#| message: false
#| warning: false

library(dplyr)
library(purrr)
library(stringr)
library(lubridate)
library(DBI)
library(duckdb)
```

## Connect to the daily trips DuckDB

```{r}
# We connect to the MITMS daily trips duckdb:
mitma_daily_con <- dbConnect(
  duckdb(),
  dbdir = "../data/mitma_daily/mitma_daily.duckdb",
  read_only = TRUE # We won't edit the database in this script
  )

# We set the maximum CPU threads and RAM usage, we are working on a powerful
# computer, these values should be adapted to your machine:
dbExecute(mitma_daily_con, "SET memory_limit = '128GB'")
dbExecute(mitma_daily_con, "SET threads = 24")

# Load the lazy table:
mitma_daily <- tbl(mitma_daily_con, "mitma_daily")

# Following Renninger and Cabrera (2026) we exclude the Canary Islands,
# Ceuta, and Melilla from our analysis, so we drop them:
mitma_daily <- mitma_daily |>
  filter(!str_starts(id_destination, "35|38|51|52"),
         !str_starts(id_origin, "35|38|51|52"))
```

## Load the UTCI table

We settled on 14:00 UTC (16:00 local time) so we keep only that hour:

```{r}
utci_per_mitma <- readRDS("../data/mitma_zones_utci_added/utci_per_mitma.rds")

utci_1400 <- utci_per_mitma |>
  select(mitma_zone_id, date, "utci" = h14)
```

As in script 06, we register the tibble as a virtual table in the DuckDB connection so that it can be joined with the daily trips database.

```{r}
# Register the tibble as a virtual table:
duckdb_register(mitma_daily_con, "utci_1400", utci_1400)

# Make it a lazy table to be able to join:
utci_db <- tbl(mitma_daily_con, "utci_1400")
```

Finally, the dates we loop over below:

```{r}
# The dates of the study period. They have to be characters, the for loop
# converts the "date" class into numbers, we need the full string to be
# able to join by date with the mitma_daily DuckDB database:
study_dates <- c(
  seq(as_date("2022-05-01"), as_date("2022-09-30")),
  seq(as_date("2023-05-01"), as_date("2023-09-30"))
  ) |>
  as.character()
```

## Aggregate the trips

We create a variable `same_district`: it is `TRUE` for the trips within a district, and `FALSE` for the trips between districts.

```{r}
daily_panels <- vector("list", length(study_dates))
names(daily_panels) <- study_dates

# Run the for loop for every day in our study period:
for (one_date in study_dates) {

  daily_panels[[one_date]] <- mitma_daily |>
    filter(date == one_date) |>
    # Join the UTCI data with mitma_daily:
    left_join(utci_db, by = join_by(id_destination == mitma_zone_id, date)) |>
    # Flows within a district are the ones whose origin and destination are
    # the same district, flows between districts are all the others:
    mutate("same_district" = if_else(id_origin == id_destination,
                                     TRUE, FALSE)) |>
    # Group by the variables the model uses, and sum the trips:
    group_by(date, id_destination, activity_destination, same_district, utci) |>
    summarise("n_trips" = sum(n_trips), .groups = "drop") |>
    collect() |>
    # DuckDB returns the ids as factors, and we want them as text in the
    # panel so that later joins do not have to convert them:
    mutate(
      "id_destination" = as.character(id_destination),
      # The province, to cluster the standard errors. It is
      # the first two digits of the district id:
      "province" = str_sub(id_destination, end = 2))
}

panel_table_s2 <- list_rbind(daily_panels)
```

The districts that have no UTCI are the Portuguese and French ones, plus the trips whose destination is outside the study area. We drop them:

```{r}
panel_table_s2 <- panel_table_s2 |>
  filter(!is.na(utci))

glimpse(panel_table_s2)
```

## Save the panel

```{r}
dir.create("../data/model_panels", showWarnings = FALSE, recursive = TRUE)

saveRDS(panel_table_s2, "../data/model_panels/panel_table_s2.rds")

dbDisconnect(mitma_daily_con, shutdown = TRUE)
```
