---
title: "The palette of the TFM"
author: "Miguel Agenjo Montejo"
format: html
editor: visual
---

```{r}
library(ggplot2)
library(RColorBrewer)
library(colorBlindness)
```

In order to have a shared style across all the plots in the masters thesis and to avoid repeating the code for this shared style, we created a custom theme that we can add to any ggplot. This documents explains the decisions taken in the code that can be checked at the accompanying file `theme_tfm.R`. We also explain the decisioins regarding the color palette.

## Font

Since we are using the APA Latex template UC3M offers, we decided to follow [APA font suggestions for figures](https://apastyle.apa.org/style-grammar-guidelines/paper-format/font): "a sans serif font with a type size between 8 and 14 points". We settled on Atkinson Hyperlegyble Next. This font, created by the Braille Institute, meets the san serif criteria and, additionaly, improves accesibility since it was ["designed to improve legibility and readability for individuals with low vision"](https://www.brailleinstitute.org/freefont/).

## Base theme

We didn't create the settings for our custom theme from scratch, we use `theme_minimal` from `ggplot2` as a base. This theme provides a clean and accesible style as a foundation, to which we added some tweaks:

- Cuando termine los añado, están en el script, igual es redundante esto y lo quito por completo.


## Color palette

We followed Wickham, Navarro, and Lin's [*ggplot2: Elegant Graphics for Data Analysis (3e)*](https://ggplot2-book.org/scales-colour.html) suggestions when deciding what color palette to use. Keeping in mind color blind users, we decided to use the following palettes:

- For *discrete variables*, we will use a `Set2`, and `Dark2`. `Dark2` is "particularly good for points" and `Set2` is among the palettes that "work well for areas" ([Wickham, Navarro, & Pedersen, 2026, Chapter 11](https://ggplot2-book.org/scales-colour.html#brewer-scales)).

```{r}
display.brewer.all(select = c("Set2", "Dark2"))

brewer.pal.info[c("Set2", "Dark2"), ]

# The package `coloBlindness` allows us to have an approximate 
# idea of how the colors would look for different kinds of colorblindness
displayAllColors(brewer.pal(8, "Set2"))
displayAllColors(brewer.pal(8, "Dark2"))
```

For *ordered variables* (whether discrete or continuous) we will use another palette from the same package: `YlGnBu`, which is also colorblind friendly. It can be transformed into a continous variable using `scale_fill_distiller()` and `scale_colour_distiller()`.

```{r}
display.brewer.all(select = "YlGnBu")

brewer.pal.info["YlGnBu", ]

# The package `coloBlindness` allows us to have an approximate 
# idea of how the colors would look for different kinds of colorblindness:
colorBlindness::displayAllColors(brewer.pal(n = 9, "YlGnBu"))
```
- When used in maps, to visibly differenciate NA values, since the default gray can be hard to see within the blue range, we use the default red at alpha = 50.

- When we want to represent discrete ordered variables not in maps (points, lines, etc.), the lighter (yellow) side of the scale is too light to work against a white background. In those cases, we will ask for more colour steps than needed and remove the lighter ones. For example, if we need three colour steps, we ask for four and remove the two darker ones:

```{r}
# We would not use the first one if we need 3 steps:
display.brewer.pal(4, "YlGnBu")
```
For plots where we we only need *one color*, we will just use black black and gray. For plots where we need to *highlight one element against the rest*, we will use the orange (second step in `Dark2`) against gray. If we need a combination of just two colors... ya lo veremos, me gusta orange y blue o yellow y blue, pero depende.

```{r}
# The orange of Dark2:
brewer.pal(8, "Dark2")[2]
```


