---
title: "The palette of the TFM"
author: "Miguel Agenjo Montejo"
format: html
editor: visual
---

In order to have a shared style across all the plots in the masters thesis and to avoid repeating the code for this shared style, we created a custom theme that we can add to any ggplot. This documents explains the decisions that determine what`theme_tfm.R` does, and the decisions regarding the color palette.

```{r}
library(ggplot2)
library(RColorBrewer)
library(colorBlindness)
```

## Font

Since we are using the APA Latex template UC3M offers, we decided to follow [APA font suggestions for figures](https://apastyle.apa.org/style-grammar-guidelines/paper-format/font): "a sans serif font with a type size between 8 and 14 points". We settled on Atkinson Hyperlegyble Next. This font, created by the Braille Institute, meets the sans serif criteria and, additionally, improves accessibility since it was ["designed to improve legibility and readability for individuals with low vision"](https://www.brailleinstitute.org/freefont/).

## Base theme

We didn't create the settings for our custom theme from scratch, we use `theme_minimal` from `ggplot2` as a base. This theme provides a clean and accessible style as a foundation, to which we added some tweaks.

## Color palette

We followed Wickham, Navarro, and Lin's [*ggplot2: Elegant Graphics for Data Analysis (3e)*](https://ggplot2-book.org/scales-colour.html) suggestions when deciding what color palette to use. Keeping in mind color blind users, we decided to use the following palettes:

- For *discrete variables*, we will use a `Set2`, and `Dark2`. `Dark2` is "particularly good for points" and `Set2` is among the palettes that "work well for areas" ([Wickham, Navarro, and Lin, 2026, Chapter 11](https://ggplot2-book.org/scales-colour.html#brewer-scales)).

```{r}
display.brewer.all(select = c("Set2", "Dark2"))

brewer.pal.info[c("Set2", "Dark2"), ]

# The package `coloBlindness` allows us to have an approximate 
# idea of how the colors would look for different kinds of colorblindness
displayAllColors(brewer.pal(8, "Set2"))
displayAllColors(brewer.pal(8, "Dark2"))
```

For *ordered variables* (whether discrete or continuous) we will use another palette from the same package: `YlGnBu`, which is also colorblind friendly. It can be transformed into a continous variable using `scale_fill_distiller()` and `scale_colour_distiller()`.

```{r}
display.brewer.all(select = "YlGnBu")

brewer.pal.info["YlGnBu", ]

# The package `coloBlindness` allows us to have an approximate 
# idea of how the colors would look for different kinds of colorblindness:
colorBlindness::displayAllColors(brewer.pal(n = 9, "YlGnBu"))
```

- When used in maps, to visibly differentiate NA values, since the default gray can be hard to see within the blue range, we use the default red at alpha = 50. No maps made it into the final report, but this decision is applied in the maps you may find if you run the scripts.

- When we want to represent discrete ordered variables not in maps (points, lines, etc.), the lighter (yellow) side of the scale is too light to work against a white background. In those cases, we ask `display.brewer.pal()` for more colour steps than needed and ignore the lighter ones. For example, if we need three colour steps, we ask for four and ignore the lighter one:

```{r}
# We would not use the first one if we need 3 steps:
display.brewer.pal(4, "YlGnBu")
```


We also used *two extra color scales for ordered variables*: when we need to represent UTCI bins we use `YlOrRd`, since it represents a color scheme traditionally associated with heat. When we need to represent age groups we use `Purples`, associating the lighter colors with the younger groups and the darker tones with the older ones. This way we make sure the reader easily realizes that color refers to a different variable. We also discard the lighter steps when necessary.

These palettes are also colorblind-friendly:

```{r}
brewer.pal.info["YlOrRd", ]
colorBlindness::displayAllColors(brewer.pal(n = 9, "YlOrRd"))
```

```{r}
brewer.pal.info["Purples", ]
colorBlindness::displayAllColors(brewer.pal(n = 9, "Purples"))
```

For plots where we we only need *one color*, we will just use black and gray when we don't need to fill large areas (for example, dots or lollipops), and blue (the third step in `Set2`) and gray when we need to fill large areas. For plots where we need to *highlight one element against the rest*, we will use the orange (second step in `Dark2`) against gray.

```{r}
# The blue from Set2:
brewer.pal(8, "Set2")[3]

# The orange from Dark2:
brewer.pal(8, "Dark2")[2]
```



